### 2018-06-30 paper/paper.tex
Paper init

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

%\documentclass[twoside,twocolumn]{article}
%\documentclass{article}[11pt]
\documentclass[fleqn]{article}

\usepackage{listings}

\lstset{
  xleftmargin=0em,
}
\usepackage{xcolor} %custom colours
\usepackage{mdframed} %nice frames

\definecolor{light-gray}{gray}{0.95} %the shade of grey that stack exchange uses

%Put in the main document

\usepackage{blindtext} % Package to generate dummy text throughout this template

\usepackage[sc]{mathpazo} % Use the Palatino font
\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
\linespread{1.05} % Line spacing - Palatino needs more space between lines
\usepackage{microtype} % Slightly tweak font spacing for aesthetics

\usepackage{amsfonts}

\usepackage{amsmath}

\usepackage[english]{babel} % Language hyphenation and typographical rules

\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry} % Document margins
\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
\usepackage{booktabs} % Horizontal rules in tables

\usepackage{lettrine} % The lettrine is the first enlarged letter at the beginning of the text

\usepackage{enumitem} % Customized lists
\setlist[itemize]{noitemsep} % Make itemize lists more compact

\usepackage{abstract} % Allows abstract customization
\renewcommand{\abstractnamefont}{\normalfont\bfseries} % Set the "Abstract" text to bold
\renewcommand{\abstracttextfont}{\normalfont\small\itshape} % Set the abstract itself to small italic text

\usepackage{titlesec} % Allows customization of titles
\renewcommand\thesection{\Roman{section}} % Roman numerals for the sections
\renewcommand\thesubsection{\roman{subsection}} % roman numerals for subsections
\titleformat{\section}[block]{\large\scshape\centering}{\thesection.}{1em}{} % Change the look of the section titles
\titleformat{\subsection}[block]{\fontsize{10}{10}\bfseries}{\thesubsection.}{1em}{} % Change the look of the section titles

\usepackage{fancyhdr} % Headers and footers
\pagestyle{fancy} % All pages have headers and footers
\fancyhead{} % Blank out the default header
\fancyfoot{} % Blank out the default footer
\fancyhead[C]{Controlled Natural Language and Robotics $\bullet$ Calvin Claus} % Custom header text
\fancyfoot[RO,LE]{\thepage} % Custom footer text

\usepackage{titling} % Customizing the title section

\usepackage{hyperref} % For hyperlinks in the PDF

%----------------------------------------------------------------------------------------
%	TITLE SECTION
%----------------------------------------------------------------------------------------

\setlength{\droptitle}{-4\baselineskip} % Move the title up

\pretitle{\begin{center}\Huge\bfseries} % Article title formatting
  \posttitle{\end{center}} % Article title closing formatting
\title{Controlled Natural Languages and Robotics} % Article title
\author{%
  \textsc{Calvin Claus (1429713)}
  \normalsize TU Wien\\ % Your institution
  \normalsize \href{mailto:e1429713@student.tuwien.ac.at}{e1429713@student.tuwien.ac.at} % Your email address
}
\date{\today} % Leave empty to omit a date
\renewcommand{\maketitlehookd}{%
  \begin{abstract}
    \noindent This paper gives a brief introduction into Controlled Natural Languages (CNLs), providing an overview over common language-rules and outlining the long term goals of CNL research. It describes how to construct a simple CNL interface to allow robots to receive human commands. Finally, it analyzes if a robot can successfully build a useful knowledge base over time, using CNL input by a human speaking naturally. It finds that, at the current state of CNL research, CNL can be used to build a good knowledge base, but the human has to unnaturally adapt his/her speaking to the fact that a robot should understand him/her, even though he/she is conforming to CNL rules.
  \end{abstract}
}

%----------------------------------------------------------------------------------------

\begin{document}

% Print the title
\maketitle

%----------------------------------------------------------------------------------------
%	ARTICLE CONTENTS
%----------------------------------------------------------------------------------------

\section{Introduction}

\lettrine[nindent=0em,lines=3]{I} t's a fascinating idea:

\section{Ivy}
Ivy is a system that is designed to verify safety of infinite-sate systems.
It aims to assist protocol architects, algorithm authors, and the like in proofing the safety and correctness of their algorithm or protocol.

\subsection{How Ivy is used}

To understand how Ivy can be used, a common workflow that incorporates Ivy in algorithm verification is presented:

\begin{enumerate}
  \item design the algorithm
    \begin{itemize}
      \item purely creative task
      \item Ivy offers no support in this area
    \end{itemize}
  \item encode it in Ivy language
    \begin{itemize}
      \item a one to one translation is most likely not possible
      \item user must be ready for the paradigm and reasoning change when writing Ivy language as opposed to a procedural programming language
    \end{itemize}
  \item verify the properties one designed the algorithm to fulfil
    \begin{itemize}
      \item specify properties as invariants
      \item make invariants inductive
    \end{itemize}
\end{enumerate}

\subsubsection{Step 1 - algorithm design}
Step 1 is executed entirely independently of Ivy. The Ivy language or the Ivy system do not provide any support in actually coming up with an algorithmic solution to a problem.
The step is mentioned here to show the context the Ivy system can be used in: to answer the question "is an algorithm correct"?

\subsubsection{Step 2 - encoding in Ivy language}
Step 2 requires the user to understand the "unusual" \cite{refLanguageDoc} - compared to a procedural programming language -  Ivy language features. The user must
find a way to use the Ivy language features to encode not only the algorithm itself but also the "system context" or "framework" in which the algorithm was designed to execute.
For example, to reason about distributed consensus algorithms Lynch has defined a "context" or "framework" in which distributed algorithms operate.
This is explained in detail later on in this work. Part of this framework describes that the network is defined as a graph, where edges can hold messages.
The system executes two distinct steps: one in which messages can be sent and another in which the states of nodes are altered.
The point is, both the algorithm and the framework in which it operates have to be flawlessly described through the Ivy language in this step. This process is
fully creative and will - judging by the authors experience - most likely not allow a mindless one-to-one translation from a procedural algorithm.
In this step the user also has to tell Ivy which methods (or "actions" in Ivy speak), e.g. \textit{connect\_to\_server(node)}, the environment can execute, . These actions are called "exported".
These are the actions the final user of the algorithm would call, and it is our desire to proof correctness of these methods. Furthermore the user can supply
the program with assertions, that they wish to proof cannot fail.

\subsubsection{Step 3 - verification}
The verification happens in step 3. The goal is to specify inductive invariants that describe the properties to be proven about the algorithm. "Inductive" is an important
keyword here. An invariant is inductive iff the following three properties are satisfied:
\begin{enumerate}
  \item initiation: the invariant is true in all legal initial states
  \item safety: for any program state that satisfies the invariant - regardless of whether this state is reachable - no exported action can cause an assertion failure starting in that state
  \item consecution: for any program state that satisfies the invariant - regardless of whether this state is reachable - then after executing any exported action, the invariant remains satisfied
\end{enumerate}
Inductive invariants is one of the more difficult concepts to get ones head around, especially as a user who does not have any experience with automatic verification.
Ivy does not consider the reachability of a system state.  As long as the state is in compliance with the invariants it is assumed to be reachable.
The author of this paper needed some time to understand why Ivy generates counterexamples that are clearly not reachable, e.g. negative values for counters that are set to zero at initialization.
To fix this one would have to make such the counter part of the invariant e.g. $invariant counter >= 0$.

The name "Ivy" stems from "interactively verifying", which highlights the key differentiator of Ivy to Coq or Dafny.
Ivy actively supports the user in the search for an inductive invariant by graphically displaying a concrete minimal counterexample to induction.
The user can then generalize from the counterexample to strengthen the invariants. This counterexample $\rightarrow$ strengthened invariant $\rightarrow$ counterexample process continues
until an inductive invariant is found and the correctness of the algorithm is proven. However, there is no mechanism in place that automatically disallows invariants that exclude reachable states.
This means that it's easy to exclude theoretically reachable edge cases one hasn't thought of, by writing invariants that are too strong.

\subsection{The Ivy language}
The Ivy language is a core part of the Ivy system. It is the language through which the algorithm to be verified is defined.
This section presents a selection of core language features that should help the reader understand the paradigm of modelling in Ivy, as well as lay the necessary foundation to understand
the rest of the work presented in this paper. This section oversimplifies and does not give an exhaustive documentation, which is given in \cite{refLanguageDoc}.

\subsubsection{Types and declarations}
In Ivy one mainly uses uninterpreted types. ``Uninterpreted'' means that nodes in the system state are an arbitrary set with at least one element. The cardinality of this
set is parameterized.  Meaning how many nodes are generated to construct a counter example to induction is entirely up to Ivy, one does not manually instantiate instances of these types.

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
type node
\end{lstlisting}
\end{mdframed}

\noindent To relate instances of types to each other we can declare functions.

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
type node
type value
function value(N:node) -> value
\end{lstlisting}
\end{mdframed}

\noindent A function that maps to bool can be declared as a relation, so

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
function message_received(RECEIVER:node, SENDER:node) -> bool
\end{lstlisting}
\end{mdframed}

\noindent is the same as

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
relation message_received(RECEIVER:node, SENDER:node)
\end{lstlisting}
\end{mdframed}

\subsubsection{Assignments}

To assign values to relations, functions or variables $:=$ is used. To assign to multiple values
at the same time one can use an uppercase placeholder. For example setting all nodes with an incoming message to failed:

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
failed(X) := ~incoming_message(X);
\end{lstlisting}
\end{mdframed}

\subsubsection{Actions}
To define a callable action, that is callable from within another action or from the environment if it is exported
this is the syntax.

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}

action fail_node_and_send(x:node) = {
  fail(x) := true;
  call send;
}
export fail_node_and_send

action send = {
  sent(X) := true;
}
\end{lstlisting}
\end{mdframed}

\subsubsection{Non deterministic choice}
To assign to a variable, relation or function a value non-deterministically (randomly) one uses ``*''.
This line non-deterministically chooses a number of nodes to have failed.

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
failed(X) := *;
\end{lstlisting}
\end{mdframed}

\subsubsection{Assume, require and ensure}
To specify safety conditions in an action we can use require and ensure.
They both act like a traditional assert in that it does not allow control to pass if it is false.
They are different semantically in that require assigns blame to the actions caller, while ensure
is a condition that the action itself must guarantee. This is especially relevant to restrict
the environment in the way it can call an action.

Assume is different in that it does not allow control to pass through if the condition is not satisfied.
The author has found this to be mostly useful in combination with non-deterministic assignments that one
wants to restrict.

The following action chooses a random node to assign to y, but will not allow control to pass if
the chose node has failed. Therefore the actions will choose a random non failing node to send a message to x.
If there is no node that hasn't failed, this action cannot terminate.

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
action random_node_sends_message_to(x:node) = {
  y := *;
  assume ~failed(y);
  send(x, y);
}
\end{lstlisting}
\end{mdframed}

Assume simplifies modelling in some situations, as it can easily eliminate impossible behaviour. However,
it is risky to use assume for the same reason: one might dismiss a valid counterexample to the specification
by assuming it cannot happen.

\subsubsection{Invariant}
To specify the properties we want to proof about our algorithm we encode them in inductive invariants like so:

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
invariant system.rounds >= 0
\end{lstlisting}
\end{mdframed}

These must hold at all times when an action is not executing, thus they have to be initially true and each of
the exported actions must preserve them. However, an invariant is not guaranteed to hold when an action calls another action.

\subsection{Decidability}
Not all first-order formulas are decidable in Ivy. Therefore, the system defines a subset of first-order formulas as its decidable fragment.
This not only constraints the structure of logical formuals but also functions. In Ivy only stratified functions are allowed. Meaning that if a graph was constructed
with every type being a node and every function from a type A mapping to another type B would be represented as an edge the graph has to be cycle free. Thus there cannot be a function from
$X \rightarrow Y$ and from $Y \rightarrow X$ in the same specification.

Ivy is extremely vocal about this and will let the user know if his specification does not lie in the decidable fragment.

\section{Distributed Consensus Algorithms with stopping failures}

The algorithm that was implemented in Ivy as part of this work is called \textit{Floodset} and was taken out of Lynch's work ``Distributed Algorithms'' \cite{refNancy}.
\textit{Floodset} is a consensus algorithm for synchronous networks that is designed to reach consensus in a network with failing processes.
To understand \textit{Floodset}, one needs to first be made familiar with the aforementioned definition of the framework/system within which this algorithm operates

\subsection{Framework - Synchronous Network Systems}

The network of processes is represented as a directed graph $G=(V,E)$. Where $n$ is $|V|$ the number of nodes in the network. Each node represents a process in the network.
The communication paths between nodes are represented by directed edges in the graph.
$out-nbrs_i$ is the set of nodes to which an edge from node $i$ exists.
$in-nbrs_i$ is the set of nodes from which an edge to node $i$ exists.
$M$ denotes the message alphabet, meaning that elements of $M$ can be used to construct messages (vectors of elements of $M$) from nodes to other nodes.
There are a number of components associated with each node:
\begin{itemize}
    \item $states_i$, a set of states (can be infinite)
    \item $start_i$, a nonempty subset of $states_i$
    \item $msgs_i$, a message generation function that maps $states_i X out-nbrs_i$ to elements of $M U {null}$
    \item $trans_i$, a state transition function that maps $states_i X vectors of M index by in-nbrs_i$ to $states_i$
\end{itemize}
Thus, a node is always in some state in $states_i$. Importantly, the set of states is not necessarily finite, meaning that unbounded data structures like arrays can be modeled in the system.
Each node defines a message generation function that uses the current state to return a message for each out neighbour. Similarly each node has a state transition function that is used to determine
from the current state and all incoming messages the state the next state the node will transition to.

The edges in the graph are called channels and can hold at most one message at a time or $null$ to denote an empty channel.
The system of nodes operates in steps. It begins with all nodes in an arbitrarily chosen start state from $start_i$ and all channels empty.
The processes then all repeatedly perform the following two steps \textit{synchronously}:
\begin{itemize}
  \item Apply the message generation function inputting the current state and one of the outgoing neighbours. Put these message in the channel to the chosen outgoing neighbour. Do this for each outgoing neighbour.
  \item Apply the state transition function to the current state and the array of incoming messages. Transition the node to the returned state. Remove all messages from the channels.
\end{itemize}
\textit{Synchronously} in this case means that all nodes wait for each other to finish step1 before executing step2 and vice versa.
The combinations of these two steps is a \textit{round}.

\subsection{Stopping failures}
The system presented in Lynch's work \cite{refNancy} is capable of modelling failure in a variety of ways.
This work makes use of stopping failure only. This kind of failure occurs when a process stops in the middle of its execution. With regards to the presented framework this
is modeled as a process failing before or after performing some instance of Step 1 or Step 2. Additionally, a process may also stop in the middle of performing Step 1. This allows
the interesting situation in which a process has communicated its messages to some but not yet all processes. It is important to note that processes are not thought of performing
the message generation operation sequentially. Instead, we say that the order in which messages are put in the channels is non-deterministic. This clarification, allows a failing process
to send its messages to an arbitrary subset of its outgoing neighbours only.

\subsection{Floodset}
The problem at hand is as follows: An arbitrary number of processes start with individual inputs from a value set $V$. This is modeled by each node having at least one start state
for every possible value $v ∈ V$. Also, a default value $v_0 ∈ V$ that is equal for all nodes is defined.
Processes communicate with each other through several \textit{rounds}. The nature of their communication and its effect on their states is defined by their $msgs_i$ and $trans_i$ functions.
Nodes may fail according to the \textit{stopping failure} model described above. The total number of failing processes is bounded by a constant $f$. The entire network knows the value of $f$ and
can therefore utilise it in their state transition function.
All processes that did not fail must eventually decide on a single value from the value set $V$. \textit{Deciding} is modeled by nodes setting a special \textit{decision} state component to a value in $V$.
Their decision is subject to the following correctness conditions:
\begin{itemize}
  \item \textbf{Agreement}: No two processes decide on different values.
  \item \textbf{Validity}: If all processes start with the same initial value $v ∈ V$, then $v$ is the only possible decision value
  \item \textbf{Termination}: All nonfaulty processes eventually decide.
\end{itemize}
We assume the processes are in a complete, undirected graph. This means that every node has complete knowledge over the network. Channels are fully reliable - every message gets delivered.

In the \textit{Floodset} algorithm processes hold a set $W$ as one of their state components. This set holds the subset of values of $V$ that the process has seen. Every node's initial $W$ holds only its initial value.
In Step 1 of every round nodes broadcast their $W$. In Step 2 all $v ∈ V$ that \textit{(a)} are sent to a process and \textit{(b)} the process does not yet hold in its $W$ are added to its $W$. In the $f+1$th round
a node decides: If $W$ is a singleton set then it decides on the unique element of $W$, otherwise it decides on the default value $v_0$.

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\noindent $states_i$:
\begin{gather*}
  rounds   \in  \mathbb{N}, \ \mathrm{initially} \ 0 \\
  decision  \in  V \cup \{ unknown \}, \ \mathrm{initially} \ unknown \\
  W \subseteq V, \ \mathrm{initially \ the\  singleton\  set\  consisting\  of\  i's\  initial\  value}
\end{gather*}
\end{mdframed}

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\noindent $msgs_i$:
\begin{gather*}
  if\ rounds \le f\ \mathrm{send}\ W\ \mathrm{to\ all\ other\ processes}
\end{gather*}
\end{mdframed}

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\noindent $trans_i$:
\begin{gather*}
  rounds\ :=\ rounds+1\\
  \mathrm{let}\ X_j\ \mathrm{be\ the\ message\ from\ j,\ for\ each\ j\ from\ which\ a\ message\ arrives}\\
  W\ :=\ W \cup \bigcup_{j} X_j\\
  \mathrm{if}\ rounds = f+1\ \mathrm{then}\\
  \ \ \   \mathrm{if}\ |W| = 0\ \mathrm{then}\ decision := v,\ \mathrm{where}\ W = {v}\\
  \ \ \   \mathrm{else}\ decision := v_0
\end{gather*}
\end{mdframed}

\noindent What follows is a proof of the correctness of the \textit{Floodset} algorithm to illustrate the principles behind it to the reader.

Let $W_i(j)$ denote the value of $W$ for node $i$ in round $j$. The following Lemmas apply:\\

\noindent \textbf{Lemma 1}:\\
If no process fails during a particular round $r,\ 1 \le r \le f + 1$, then $W_i(r) = W_j(r)$ for all $i$ and $j$ that did not fail after $r$ rounds.\\
\textbf{Proof}:\\
\textit{Indirect}: If there exists a value $v \in V$ that is in $W_i(r)$ but not in $W_j(r)$ after round $r$ in which no process failed, that implies that some node $k$ sent a value to $i$ that it did not send to $j$, or $i$ failed to send a member of $W_i(r-1)$. Since all nodes broadcast their $W$, this is only possible if node $k$ or $i$ failed to send at least one message. But since no processes failed, each process must have shared their $W$ with every other process. \textit{Contradiction}. \\

\noindent \textbf{Lemma 2}:\\
Suppose that $W_i(r) = W_j(r)$ for all $i$ and $j$ that did not fail after $r$ rounds. Then for any round $r',\ r \le r' ≤ f + 1$, the same holds, that is, $W_i(r’) = W_j(r’)$ for all $i$ and $j$ that did not fail after $r’$ rounds.\\
\textbf{Proof}:\\
Since all active nodes after $r$ rounds have equal $W$s there exists no $v \in V$ that is in some $W_i(r),\ i \mathrm{some\ active\ node}$ that is not in every other $W_j(r),\ j \mathrm{any\ active\ node}$.
Since nodes will only send their $W$s, this implies that no node can send a message in some future round $r'$ that contains a value $v$ that some node does not have in its $W$, thus $W_i(r’) = W_j(r’)$, $i,\ j$ not failed.\\

\noindent \textbf{Lemma 3}:\\
After $f+1$ rounds there has been a round $r'$ in which no process failed\\
\textbf{Proof}:\\
\textit{Indirect:} If a process failed in all of $f+1$ rounds at least on process must have failed in all of the $f+1$ rounds. This implies
that at least $f+1$ processes failed. Only $f$ process can fail. \textit{Contradiction}.\\

\noindent \textbf{Lemma 4}:\\
If processes $i$ and $j$ are both active after $f+1$ rounds, then $W_i = W_j$ at the end of round $f + 1$.\\
\textbf{Proof}:\\
It follows from \textit{Lemma 3}, that after $f+1$ rounds at least one round, $r'$, in which no process failed, occurred. From \textit{Lemma 1} we therefore derive that $W_i(r') = W_j(r')$ for all active $i,\ j$.
From \textit{Lemma 2} it follows that $W_i(f+1) = W_j(f+1)$. \textit{Q.e.d}.\\

\noindent The Lemmas allow proofing the theorem:\\
\noindent \textbf{Thorem}:\\
\textit{Floodset} satisfies the correctness conditions.\\
\textbf{Agreement}:\\
From \textit{Lemma 4} we know what $W_i = W_j$ after $f+1$ rounds. Since all decision conditions are the same and only depend on $W$ agreement is satisfied.\\
\textbf{Validity}:\\
If all processes start with the same initial value $v ∈ V$, then there exists no value $v'$ that is part of any $W$. Thus, from $W_i(f+1) = W_j(f+1)$, $i,\ j$ active and $W_i(f+1) = \{v\}$ it follows
that every active node decides $v$.\\
\textbf{Termination}:\\
Trivially, nodes decide after a fixed number of rounds, $f+1$.\\

This work now presents a way to translate Lynch's framework and the \textit{Floodset} algorithm to a specification in the Ivy language.

\section{Specification}

\section{Evaluation}

\section{Conclusion}
This paper gave a brief introduction into Controlled Natural Languages (CNLs)

%----------------------------------------------------------------------------------------
%	REFERENCE LIST
%----------------------------------------------------------------------------------------
\begin{thebibliography}{99}

    \bibitem{refLanguageDoc} http://microsoft.github.io/ivy/language.html
    \bibitem{refNancy} http://microsoft.github.io/ivy/language.html
\end{thebibliography}

%----------------------------------------------------------------------------------------

\end{document}

### 2018-07-05 paper/paper.tex
Paper progress

@@
 \textbf{Proof}:\\
-Since all active nodes after $r$ rounds have equal $W$s there exists no $v \in V$ that is in some $W_i(r),\ i \mathrm{some\ active\ node}$ that is not in every other $W_j(r),\ j \mathrm{any\ active\ node}$.
+Since all active nodes  have equal $W$s after $r$ rounds, there exists no $v \in V$ that is in some $W_i(r),\ i\ \mathrm{some\ active\ node}$ that is not in every other $W_j(r),\ j\ \mathrm{any\ active\ node}$.
 Since nodes will only send their $W$s, this implies that no node can send a message in some future round $r'$ that contains a value $v$ that some node does not have in its $W$, thus $W_i(r’) = W_j(r’)$, $i,\ j$ not failed.\\
@@
 \textbf{Agreement}:\\
-From \textit{Lemma 4} we know what $W_i = W_j$ after $f+1$ rounds. Since all decision conditions are the same and only depend on $W$ agreement is satisfied.\\
+From \textit{Lemma 4} we know what $W_i = W_j$ after $f+1$ rounds. Since all decision conditions are the same and only depend on $W$, agreement is satisfied.\\
 \textbf{Validity}:\\
-If all processes start with the same initial value $v ∈ V$, then there exists no value $v'$ that is part of any $W$. Thus, from $W_i(f+1) = W_j(f+1)$, $i,\ j$ active and $W_i(f+1) = \{v\}$ it follows
+If all processes start with the same initial value $v \in V$, then there exists no value $v', v \neq v'$, that is part of any $W$.
+Thus, from $W_i(f+1) = W_j(f+1)$, $i,\ j$ active, and $W_i(f+1) = \{v\}$ it follows
 that every active node decides $v$.\\
 \textbf{Termination}:\\
-Trivially, nodes decide after a fixed number of rounds, $f+1$.\\
+Trivially, nodes decide after $f+1$ rounds.\\
 
@@
 \section{Specification}
+To summarize: at hand we have a parameterized system with $n$ process and $f$ number of faults.
+We want to show \textit{Floodset} is correct regardless of the parameters.
+It is trivial to verify for a small number of $n,\ f$, but the question is ``does \textit{Floodset} work for any assignment of $n,\ f, n \lt f$''.
+To answer this question we are using the theorem proofer Ivy; first encoding the algorithm in the Ivy language and then constructing inductive invariants to proof the correctness.
 

### 2018-07-07 paper/paper.tex
paper progress

@@
 \usepackage{hyperref} % For hyperlinks in the PDF
+
+\lstset{basicstyle=\small}
 
@@
 We want to show \textit{Floodset} is correct regardless of the parameters.
-It is trivial to verify for a small number of $n,\ f$, but the question is ``does \textit{Floodset} work for any assignment of $n,\ f, n \lt f$''.
+It is trivial to verify for a small number of $n,\ f$, but the question is ``does \textit{Floodset} work for any assignment of $n,\ f,\ n < f$''.
 To answer this question we are using the theorem proofer Ivy; first encoding the algorithm in the Ivy language and then constructing inductive invariants to proof the correctness.
 
+The goal of this section is to walk through the specification step by step, to give the reader an understanding of the paradigms relevant to automatic verification and Ivy in particular.
+The complete code can be found here \cite{github}.
+
+\subsection{Modeling strategy}
+
+\subsubsection{Modularity}
+When verifying an algorithm like \textit{Floodset} in Ivy, an algorithm designer would most likely prefer to be able to translate their original algorithm to the Ivy language in as close to a one-to-one fashion as possible.
+However, this is not quite convenient for \textit{Floodset} and the framework it operates in.
+To separate system and nodes one would have to rely heavily on iterations. E.g. for \textit{Step 1} the system would iterate over all nodes and let each node take care of sending it's own message.
+
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+action round = {
+	var iter := node.t.iter.create(0);
+  while ~iter.is_end
+	{
+ 	 call send_messages(iter.val)
+	 iter := iter.next;
+	}
+}
+action send_messages(n:node.t) = {
+  node.message(n, V) := node.w(n, V);
+  node.incomming_message(O, n) := true;
+}
+
+export round
+\end{lstlisting}
+\end{mdframed}
+
+However, the code above does not actually execute an iteration over all nodes as one would expect. Looking at an Ivy trace we see what really happens:
+
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+message := *;
+incomming_message := *;
+\end{lstlisting}
+\end{mdframed}
+
+Ivy non-deterministically assigns to all relations involved in the loop, instead of actually executing the loop.
+This is because proofing correctness of loops can only be done through inductive loop invariants specified by a human.
+Ivy therefore allows specifying invariants together with a loop.
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+  while ~iter.is_end
+  invariant node.message(N,V) = node.w(N,V)
+  invariant N<=iter.val -> incomming_message(O, N)
+	{
+    . . .
+	}
+\end{lstlisting}
+\end{mdframed}
+
+These invariants must hold before the loop and after every iteration.
+The following is a trace of the execution of the loop with the specified invariants.
+First they are asserted, then the relations are set to non-deterministic values and finally an assume is used to restrict control to values for the relation that fulfil the invariants.
+
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+assert message(N,V) = w(N,V)
+assert N:t <= t.iter.val(loc:iter) -> incomming_message(O,N)
+incomming_message := *
+message := *
+assume message(N,V) = w(N,V)
+assume N:t <= t.iter.val(loc:iter) -> incomming_message(O,N)
+\end{lstlisting}
+\end{mdframed}
+
+As a result of this architecture the program logic has to be specified twice: once in the form of invariants and another time in the actual loop.
+Such a program therefore has two possible origins for any error: a fault in the algorithm or an inadequate invariant.
+This structure also increases the difficulty of debugging, because Ivy only shows unhelpful $assert$, $assume$ and  non-determinsitic assignments ($:= *$) in its traces, rather than execution of the actual lines in the $send_messages$ function.
+
+For the simple \textit{Step 1} this approach may still be considered viable, but for the more complex \textit{Step 2} and the failure simulation logic, the author has found that the duplication of logic and the decreased debugging efficiency makes the increased modularity achieved through iterations not worth the mental and maintenance effort that the loop invariants bring with them.
+
+For this reason the following specification of \textit{Floodset} uses a non-modular structure where single lines predominately alter the state of the system as a whole.
+
+\subsubsection{Cardinality relationships}
+There is no cardinality operator in IVy.
+In set notation we would write something like $|\{n|node(n) \land failed(n)\}|$ to count the number of failed nodes in a system.
+When programming in a standard procedural language we would use a simple loop for this task but, as has been explained in the previous section, loops are somewhat problematic in Ivy.
+The author couldn't find a way to specify an inductive loop invariant for a loop counting the number of failed nodes without a cardinality operator at hand.
+This means that alternative encodings for the assumption $f<n$ and the statement $if rounds = f+1$ must be found.
+
+The assumption $f<n$, ``number of failed nodes is bounded to an amount smaller than the number of total nodes'', can alternatively be expressed via $\exists n\ \neg failed(N)$.
+
+To figure out how to alternatively encode $if rounds = f+1$ deep understanding of the algorithm is necessary.
+One must realise that $rounds = f+1$ entails that a ``clean round'' occurred (see \textit{Lemma 3}).
+Where ``clean round'' describes ``a round in which no node failed''.
+The processes are ready to decide after a clean round has occurred.
+They can decide either immediately, or an arbitrary number of rounds after the clean round.
+This can be modeled with non-determinism:
+
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+clean_round_occured := (clean_round_occured | (forall N . ~node.failed_during_round(N)));
+if clean_round_occured {
+  decision_round := *;
+};
+\end{lstlisting}
+\end{mdframed}
+
+And an invariant that describes the properties of a clean round, that contains the premise of \textit{Lemma 1}:
+
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+# lemma 1:
+invariant system.clean_round_occured -> ((~node.failed(A) & ~node.failed(B)) -> (node.w(A, V) = node.w(B, V)))
+\end{lstlisting}
+\end{mdframed}
+
+The (other) chosen invariants will be explained in detail later on in this paper.
+
+\subsection{Types, functions, relations}
+First the types, functions, relations and individual variables that are used throughout the program are presented:
+
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+type node
+type value
+object node = {
+  #graph/system related
+  function val(X:node) : value
+  function default(X: node) : value
+  individual def : value
+  relation failed(A:node)
+  relation failed_before_round(A:node)
+  relation failed_during_round(A:node)
+  relation has_single_value(N:node)
+
+  #state
+  relation w(SELF:node, V:value)
+  relation decision(SELF:node, VALUE:value)
+
+  #communication
+  relation message(N:node, V:value)
+  relation incomming_message(SELF:node, OTHER:node)
+  relation incomming_message_failing(SELF:node, OTHER:node)
+}
+
+object system = {
+  individual decision_round:bool
+  individual clean_round_occured:bool
+}
+\end{lstlisting}
+\end{mdframed}
+
+Their meaning will become clear as the code is presented in the following paragraphs.
+
+A round in the specification is defined as follows:
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+action round = {
+  require ~node.decision(N,V) & ~decision_round;
+  node.failed_before_round(N) := node.failed(N);
+  call step1;
+  call simulate_failure;
+  clean_round_occured := (clean_round_occured | (forall N . ~node.failed_during_round(N)));
+  if clean_round_occured {
+    decision_round := *;
+  };
+  call step2;
+}
+export round;
+\end{lstlisting}
+\end{mdframed}
+
+Line 2 makes it illegal for the environment to call \textit{round} if decisions already occurred.
+Line 3 does some bookkeeping.
+\textit{call step1} models the communication step. \textit{simulate\_failure} simulates failure before, during
+or after \textit{step1}. After \textit{decision\_round} has been set as explained above, \textit{call step2} is
+executed to model the state transition of the nodes.
+
+\subsection{Step 1 - Message sending}
+The contents of a nodes message is modeled via the relation $message(N:node, V:value)$. Where
+$message(1, 0) = true; message(1, 1) = true;$ means node $1$ includes in its message the values $0$ and $1$.
+Since a node never sends the same message to all nodes for any given round we don't need to include the
+recipient in the relations.
+
+The actual delivery of a message is modeled via the $incomming_message(RECIPIENT:node, SENDER:node)$ relation.
+Where $incoming_message(0, 1) = true; incoming_message(0, 2) = true;$ means that node $0$ receives the messages from node $1$ and node $2$.
+
+These relations allow us to use Ivy's assignments to easily model \textit{Step 1} in a round of \textit{Floodset}.
+Line 2 assigns for all nodes $A$ and values $V$ to $message(A,V)$ the boolean value of $node.w(A,V) \land \neg failed(A)$.
+In other words every node that hasn't failed prepares a message with all values in their $w$.
+Line 3 states that all nodes that haven't failed send messages to all other nodes.
+
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+action step1 = {
+  node.message(A, V) := node.w(A, V) & ~node.failed(A);
+  node.incomming_message(O, A) := ~node.failed(A);
+}
+\end{lstlisting}
+\end{mdframed}
+
+\subsection{Step 2 - Message sending}
+
+Line 2 of the \textit{step2} action adds to $w$ of node $N$ all values $V$ for which a node $J$ exists that has sent a message to $N$ that contains $J$.
+The conditional statement in Line 4 corresponds to $if\ rounds = f+1$ in the original algorithm. Nodes with a single value in $w$ decide on that value, otherwise
+they decide on the default value.
+Finally, for good measure, $incomming\_message$ and $message$ are reset, but this is not necessary for the proof.
+
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=10, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+action step2 = {
+  node.w(N,V) :=
+    node.w(N,V) | (exists J . node.incomming_message(N,J) &
+                              node.message(J,V));
+
+  if decision_round {
+
+    node.has_single_value(N) :=
+      (exists J . node.w(N,J)) &
+      (forall V,O . O=V | (node.w(N,V) -> ~node.w(N,O)));
+
+    node.decision(N,K) :=
+      node.has_single_value(N) & node.w(N,K) & ~node.failed(N);
+
+    node.decision(N,node.default(N)) :=
+      (~node.has_single_value(N) | node.decision(N,node.default(N)))
+      & ~node.failed(N);
+
+  };
+
+  node.incomming_message(A,B) := false;
+  node.message(A,V) := false;
+}
+\end{lstlisting}
+\end{mdframed}
+
+\subsection{Failure simulation}
+
+The \textit{simulate\_failure} action simulates the failure model described by Lynch.
+It is called after \textit{step1} and illustrates sophisticated usage of the $assume$ functionality in Ivy.
+In line 2 it choses a random number of nodes to fail. The assume in line 3 however excludes all
+cases where a node that has already failed was chosen to fail in this round.
+Line 4 prepares for each node $M$ a non-deterministically chosen set of nodes that it sends a message to.
+Line 6 updates the \textit{incomming\_message} relation. If node $M$ failed during this round, the set of nodes
+ $M$ sends messages to is changed to the values of \textit{incomming\_message\_failing}, otherwise it remains the same.
+ After the nodes that \textit{failed\_during\_round} are set to \textit{failed} in line 12, line 13 assumes
+ the condition $f<n$. This excludes the cases where all nodes failed.
+
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=5, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+  action simulate_failure = {
+    node.failed_during_round(N) := *;
+    assume node.failed_during_round(N) -> ~node.failed(N);
+    node.incomming_message_failing(N,M) := *;
+
+    node.incomming_message(N,M) :=
+      (node.failed_during_round(M) &
+        node.incomming_message_failing(N,M)) |
+      (~node.failed_during_round(M) &
+        node.incomming_message(N,M));
+
+    node.failed(N) := (node.failed(N) | node.failed_during_round(N));
+    assume exists N . ~node.failed(N);
+  }
+\end{lstlisting}
+\end{mdframed}
+
+Finally, we describe the initial system state in a special \textit{after init} action. When
+proofing correctness Ivy requires the initialisation to entail the invariants. All relations and functions
+that are not assigned any values here are set automatically by the environment. E.g. $node.val(N)$ is
+treated as a parameter the environment sets.
+
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=5, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+after init {
+  def := *;
+  default(N) := def;
+  w(N, V) := V=node.val(N);
+  incomming_message(A, B) := false;
+  decision(N, V) := false;
+  clean_round_occured := false;
+  decision_round := false;
+  node.failed_before_round(N) := false;
+  node.failed(N) := false;
+  node.has_single_value(N) := false;
+}
+\end{lstlisting}
+\end{mdframed}
+
+\subsection{Invariants}
+
+Now that the system has been modeled the next step of verifying correctness with Ivy is specifying inductive invariants.
+The simplest invariants to come up with are those that describe the correctness of the algorithm at hand.
+In our case these are \textit{agreement}, \textit{validity}, \textit{termination}.
+
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=5, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+####### Correctness Conditions #######
+
+# Agreement: No two processes decide on different values.
+invariant (~node.failed(A) & ~node.failed(B)) ->
+            (node.decision(A,V) = node.decision(B,V))
+
+# every node decides on at most one value
+invariant node.decision(N,V) -> (~exists O . node.decision(N,O) & O~=V)
+
+# Validity: If all processes start with the same initial value v,
+#           then v is the only possible decision value
+invariant (exists N,M . node.val(N) ~= node.val(M)) |
+          (system.decision_round ->
+            (~node.failed(N2) -> node.decision(N2,node.val(N2))))
+
+# Termination: All nonfaulty processes eventually decide.
+invariant system.decision_round <->
+            (forall N . (~node.failed(N) -> (exists V . node.decision(N,V))))
+\end{lstlisting}
+\end{mdframed}
+
+If one were to run $ivy_check$ on the program without specifying any
+additional invariants Ivy finds that the \textit{agreement} invariant is not inductive
+and produces the a counter example to induction (CTI). One can display the CTI in a GUI.
+In the left hand section of the program we can see that the system transition from state 0
+to state 1 by executing \textit{round}. In the center Ivy displays state 0, meaning the state
+before \textit{round} was executed. The brown edges specify the relation $w$, the purple edges
+show $val$.
+
+---SCREENSHOT 1 TODO---
+
+By examining state 0 we realise that no node has their initial value modeled by $val$
+in its $w$. The programmer must realise this is an unreachable state and exclude it by coming
+up with an inductive invariant. The process of finding invariants interactively is supported by Ivy but
+beyond the scope of this paper.
+
+An invariant excluding the unreachable state is one that states that the initial value of a node is always in its $w$:
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=5, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+# initial value is in W
+invariant node.w(N, node.val(N))
+\end{lstlisting}
+\end{mdframed}
+
+The method for finding the remaining invariants is identical:
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=5, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+# the global default is the one the nodes know
+invariant node.default(N) = node.def
+# if a value is in a process' W, it must be initial value of some other process ( no invented values )
+invariant node.w(N, V) -> (exists M . node.val(M)=V)
+# lemma 1:
+invariant system.clean_round_occured -> ((~node.failed(A) & ~node.failed(B)) -> (node.w(A, V) = node.w(B, V)))
+\end{lstlisting}
+\end{mdframed}
+
+Since it is also required that the framework behaves correctly it makes sense to include some
+correctness conditions specifically aimed at the system.
+
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=5, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+####### system correctness checks #######
+
+# f < n
+invariant exists N . ~node.failed(N)
+
+# 'after' a node has failed it cannot start working again
+invariant node.failed_before_round(N) -> node.failed(N)
+
+# 'after' a node has failed it will not send anything
+invariant node.failed_before_round(N) -> ~node.incomming_message(M, N)
+
+# If the decision round occured, a clean round must have occured
+invariant system.decision_round -> system.clean_round_occured
+
+\end{lstlisting}
+\end{mdframed}
+
 \section{Evaluation}
+
+How viable is Ivy as a tool for algorithm designers to proof correctness
+of their work? The goal when turning to a tool like Ivy is certainty of
+correctness. But absolute certainty is never really achieved. When specifying the
+algorithm in the Ivy language one can not trivially determine that the specification
+is logically equivalent to the actual algorithm. This is compounded by the fact
+that the Ivy language is structurally unlike classical procedural languages.
+Making it challenging to achieve a one-to-one translation from a classical distributed algorithm in a language like C.
+An example of this is the relatively compley quantifier-logic in assignments that update the global system state.
+A similar thread is posed by $assume$ statements.
+The author of this paper originally modeled the line
+
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=5, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+node.w(N,V) := (node.w(N,V) | (~node.failed(N) & exists J . (node.incomming_message(N, J) & node.message(J, V))));
+\end{lstlisting}
+\end{mdframed}
+
+as
+
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=5, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+ node.tmp_w(A,V) := node.w(A, V);
+    node.w(A, V) := *;
+    assume (node.tmp_w(A, V) | ((node.incomming_message(A, J) & node.message(J, V)))) <-> node.w(A, V);
+\end{lstlisting}
+\end{mdframed}
+ and did not discover that the $assume$ is bogus and excludes important possibilities from being verified.
+The error only became apparent when it was discovered that correctness could be proven in the absence of an invariant that should have been absolutely necessary for induction.
+
+- an invariant can exclude too much
+--todo finde so eine invariante --
+
+Furthermore, it has already been shown that the lack of easy iteration over types makes it difficult to achieve modularity.
+The lack of which introduces additional uncertainty about correct modelling, especially in distributed systems.
+
+However, as has been pointed out, it is not just the algorithm that must be correctly specified, but the underlying framework is equally important to get right.
+In the presented specification it is possible to remove the \textit{simulate\_failure} action and still receive a positive check from the Ivy system.
+This is unsurprising, as we do not have any mechanism in place that checks that the system correctly models the possibility for node failure.
+E.g.
+to check that the simulate failure action allows a node to fail and send a message only to a subset of other nodes one can purposefully insert an assertion that only fails when the desired behaviour is possible, e.g.
+insert
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=5, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+ensure ~(node.failed(N) &
+        (exists J . ~incoming_message(J, N) &
+          exists K . incomming_message(K, N) & J~=N & K~=J))
+\end{lstlisting}
+\end{mdframed}
+after the \textit{simulate\_failure} action.
+If Ivy does not find a counterexample to this assert, the specification is erroneous.
+In light of this, it would be beneficial to have a built in mechanism to verify the possibility of something to occur.
+Allowing one to conveniently verify that the modeled system allows for the possibility of a node to fail and only send messages to a subset of processes as part of the final proof.
+
+Of course a variety of bugs and exploits do not stem from a logically unsound algorithmic foundation but implementation errors.
+A wonderful example is the Heartbleed exploit in OpenSSL.
+It would have been no help to verify the simple Heartbleed protocol in Ivy before implementing it in C, since the error was in an implementation detail allowing a buffer underflow.
+The point is, the implementation of the algorithm must be tested through unit testing or bounded model checkers, despite successful verification in Ivy.
+
+- induction is quite an unintuitive for beginners to get their head around
 

### 2018-07-09 paper/paper.tex
Conclusion, fixes

--- 2018-07-05 paper/paper.tex
@@
 \documentclass[fleqn]{article}
-
+\usepackage{graphicx}
 \usepackage{listings}
@@
 \fancyfoot{} % Blank out the default footer
-\fancyhead[C]{Controlled Natural Language and Robotics $\bullet$ Calvin Claus} % Custom header text
+\fancyhead[C]{Verifying synchronous distributed algorithms using IVy $\bullet$ Calvin Claus} % Custom header text
 \fancyfoot[RO,LE]{\thepage} % Custom footer text
@@
   \posttitle{\end{center}} % Article title closing formatting
-\title{Controlled Natural Languages and Robotics} % Article title
+\title{Verifying synchronous distributed algorithms using IVy} % Article title
 \author{%
@@
   \begin{abstract}
-    \noindent This paper gives a brief introduction into Controlled Natural Languages (CNLs), providing an overview over common language-rules and outlining the long term goals of CNL research.
-It describes how to construct a simple CNL interface to allow robots to receive human commands.
-Finally, it analyzes if a robot can successfully build a useful knowledge base over time, using CNL input by a human speaking naturally.
-It finds that, at the current state of CNL research, CNL can be used to build a good knowledge base, but the human has to unnaturally adapt his/her speaking to the fact that a robot should understand him/her, even though he/she is conforming to CNL rules.
+    \noindent  This paper gives an introduction into IVy and the IVy language, explaining
+     the problem it solves, its basic principles and the usual context in which it can be utilized.
+     It presents a specification of the \textit{Floodset} algorithm to illustrate by example how
+     IVy can be employed by protocol designers to verify consensus algorithms, and the problems one might encouter.
+     Finally, the work evaluates on IVy, pointing out strengths and important limitations of automatic verification with the system.
   \end{abstract}
@@
 
-\lettrine[nindent=0em,lines=3]{I} t's a fascinating idea:
+\lettrine[nindent=0em,lines=3]{I} t's a fascinating idea: taking any (distributed) algorithm and feeding it to
+some powerful tool that evaluates it's correctness automatically. However, the problem of verifying infinite state systems
+has not been effectively automated. Most of the systems currently in place require a user to annotate the
+specification with inductive invariants. IVy does not solve this problem but instead increases user friendliness.
 
-\section{Ivy}
-Ivy is a system that is designed to verify safety of infinite-sate systems.
-It aims to assist protocol architects, algorithm authors, and the like in proofing the safety and correctness of their algorithm or protocol.
+IVy's underlying hypothesis is that ``automated methods are difficult to apply in practice not primarily because they are unreliable, but because they are opaque''; meaning that they fail in ways inexplicable to the user.
+In light of this, the IVy system was designed to display proof-failure comprehensively to the human user, making the root cause of errors more visible.
+\cite{ivy}
 
-\subsection{How Ivy is used}
+This paper will give an introduction to the IVy system and its language.
+It will explain the paradigms underlying the automatic verification tool, its language and the context in which it may be employed.
 
-To understand how Ivy can be used, a common workflow that incorporates Ivy in algorithm verification is presented:
+\textit{Floodset}, a simple distributed algorithm, will be explained and used as an example to illustrate the strengths and limitations of the IVy system in verifying algorithms operating in a network.
+The goal of this work is to give readers a foundational understanding of IVy and a head start in verifying their own distributed algorithms.
+
+This paper does not aim to be a comprehensive documentation of IVy or quantitatively compare IVy's effectiveness to other verification tools.
+
+\section{IVy}
+IVy is a system that is designed to verify correctness of infinite-sate systems.
+An algorithm or protocol is \textit{correct} or \textit{safe} if it cannot be called in any way that causes some pre-defined properties (assertions, invariants) not to be satisfied.
+IVy aims to assist protocol architects, algorithm authors, etc. in automatically proving the correctness of their algorithm or protocol.
+
+\subsection{How IVy is used}
+
+A common protocol or algorithm creation workflow, which incorporates IVy for verification, is a three step process:
 
 \begin{enumerate}
-  \item design the algorithm
-    \begin{itemize}
-      \item purely creative task
-      \item Ivy offers no support in this area
-    \end{itemize}
-  \item encode it in Ivy language
-    \begin{itemize}
-      \item a one to one translation is most likely not possible
-      \item user must be ready for the paradigm and reasoning change when writing Ivy language as opposed to a procedural programming language
-    \end{itemize}
-  \item verify the properties one designed the algorithm to fulfil
-    \begin{itemize}
-      \item specify properties as invariants
-      \item make invariants inductive
-    \end{itemize}
+  \item algorithm or protocol design
+  \item encoding/specification in IVy language
+  \item correctness and safety verification
 \end{enumerate}
+%\begin{itemize}
+%  \item purely creative task
+%  \item IVy offers no support here
+%\end{itemize}
+%\begin{itemize}
+%  \item a one to one translation is most likely not possible
+%  \item user must be ready for the paradigm change when writing IVy language as opposed to a procedural programming language
+%\end{itemize}
+%\begin{itemize}
+%  \item specify correctness and safety conditions as invariants
+%  \item make invariants inductive
+%\end{itemize}
 
 \subsubsection{Step 1 - algorithm design}
-Step 1 is executed entirely independently of Ivy. The Ivy language or the Ivy system do not provide any support in actually coming up with an algorithmic solution to a problem.
-The step is mentioned here to show the context the Ivy system can be used in: to answer the question "is an algorithm correct"?
+Step 1 is executed entirely independently of IVy. The IVy language or the IVy system do not provide any support in actually coming up with an algorithmic solution to a problem.
+The step is mentioned here to show the context the IVy system can be used in: to answer the question "is an algorithm correct"?
 
-\subsubsection{Step 2 - encoding in Ivy language}
-Step 2 requires the user to understand the "unusual" \cite{refLanguageDoc} - compared to a procedural programming language -  Ivy language features. The user must
-find a way to use the Ivy language features to encode not only the algorithm itself but also the "system context" or "framework" in which the algorithm was designed to execute.
-For example, to reason about distributed consensus algorithms Lynch has defined a "context" or "framework" in which distributed algorithms operate.
-This is explained in detail later on in this work. Part of this framework describes that the network is defined as a graph, where edges can hold messages.
+\subsubsection{Step 2 - encoding in IVy language}
+Step 2 requires the user to understand the "unusual" \cite{refLanguageDoc} - compared to a procedural programming language -  IVy language features. The user must
+find a way to use the IVy language features to encode not only the algorithm itself but also the \textit{system context} or \textit{framework} in which the algorithm was designed to execute.
+
+For example, to reason about distributed consensus algorithms Lynch has defined a \textit{framework} in which distributed algorithms operate.
+This is explained in detail later on in this work. Part of this \textit{framework} describes that the network is defined as a graph, where edges can hold messages.
 The system executes two distinct steps: one in which messages can be sent and another in which the states of nodes are altered.
-The point is, both the algorithm and the framework in which it operates have to be flawlessly described through the Ivy language in this step. This process is
-fully creative and will - judging by the authors experience - most likely not allow a mindless one-to-one translation from a procedural algorithm.
-In this step the user also has to tell Ivy which methods (or "actions" in Ivy speak), e.g. \textit{connect\_to\_server(node)}, the environment can execute, . These actions are called "exported".
-These are the actions the final user of the algorithm would call, and it is our desire to proof correctness of these methods. Furthermore the user can supply
-the program with assertions, that they wish to proof cannot fail.
+The point is, both the algorithm and the \textit{framework} within which it operates have to be flawlessly described through the IVy language in this step.
 
+This process is fully creative and will most likely not allow a mindless one-to-one translation from a procedural algorithm.
+The user also has to tell IVy which methods (\textit{actions}) the environment can execute, e.g. \textit{connect\_to\_server(node)}. These actions are called \textit{exported}.
+These are the actions and end-user of the algorithm would call. It is our desire to proof correctness of these actions.
 \subsubsection{Step 3 - verification}
-The verification happens in step 3. The goal is to specify inductive invariants that describe the properties to be proven about the algorithm. "Inductive" is an important
-keyword here. An invariant is inductive iff the following three properties are satisfied:
+The verification happens in step 3. The goal is to specify inductive invariants that describe the properties to be proven about the algorithm. \textit{Inductive} is an important
+keyword here. An invariant is \textit{inductive} iff the following three properties are satisfied:
 \begin{enumerate}
   \item initiation: the invariant is true in all legal initial states
-  \item safety: for any program state that satisfies the invariant - regardless of whether this state is reachable - no exported action can cause an assertion failure starting in that state
-  \item consecution: for any program state that satisfies the invariant - regardless of whether this state is reachable - then after executing any exported action, the invariant remains satisfied
+  \item safety: for any program state that satisfies the invariant - regardless of whether this state is reachable - no \textit{exported} action can cause an invariant failure starting in that state
+  \item consecution: for any program state that satisfies the invariant - regardless of whether this state is reachable - after executing any \textit{exported} action, the invariant remains satisfied
 \end{enumerate}
-Inductive invariants is one of the more difficult concepts to get ones head around, especially as a user who does not have any experience with automatic verification.
-Ivy does not consider the reachability of a system state.  As long as the state is in compliance with the invariants it is assumed to be reachable.
-The author of this paper needed some time to understand why Ivy generates counterexamples that are clearly not reachable, e.g. negative values for counters that are set to zero at initialization.
-To fix this one would have to make such the counter part of the invariant e.g. $invariant counter >= 0$.
+As a user without any experience in automatic verification, \textit{inductive} invariants are a difficult concept.
+IVy does not consider the reachability of a system state.  As long as the state is in compliance with the invariants it is assumed to be reachable.
+The author of this paper needed some time to understand why IVy generates counterexamples that are clearly not reachable, e.g. negative values for counters that are set to zero at initialization.
+Every counter needs an invariant, e.g. $invariant\ counter \geq 0$.
 
-The name "Ivy" stems from "interactively verifying", which highlights the key differentiator of Ivy to Coq or Dafny.
-Ivy actively supports the user in the search for an inductive invariant by graphically displaying a concrete minimal counterexample to induction.
+The name "IVy" stems from "interactively verifying", which highlights the key differentiator of IVy to Coq or Dafny.
+IVy actively supports the user in the search for an inductive invariant by graphically displaying a concrete minimal counterexample to induction (CTI).
 The user can then generalize from the counterexample to strengthen the invariants. This counterexample $\rightarrow$ strengthened invariant $\rightarrow$ counterexample process continues
-until an inductive invariant is found and the correctness of the algorithm is proven. However, there is no mechanism in place that automatically disallows invariants that exclude reachable states.
-This means that it's easy to exclude theoretically reachable edge cases one hasn't thought of, by writing invariants that are too strong.
+until an inductive invariant is found and the correctness of the algorithm is proven.
 
-\subsection{The Ivy language}
-The Ivy language is a core part of the Ivy system. It is the language through which the algorithm to be verified is defined.
-This section presents a selection of core language features that should help the reader understand the paradigm of modelling in Ivy, as well as lay the necessary foundation to understand
+\subsection{The IVy language}
+The IVy language is a core part of the IVy system. It is the language through which the algorithm to be verified is defined.
+This section presents a selection of core language features that should help the reader understand the paradigm of modelling in IVy, as well as lay the necessary foundation to understand
 the rest of the work presented in this paper. This section oversimplifies and does not give an exhaustive documentation, which is given in \cite{refLanguageDoc}.
@@
 \subsubsection{Types and declarations}
-In Ivy one mainly uses uninterpreted types. ``Uninterpreted'' means that nodes in the system state are an arbitrary set with at least one element. The cardinality of this
-set is parameterized.  Meaning how many nodes are generated to construct a counter example to induction is entirely up to Ivy, one does not manually instantiate instances of these types.
+In IVy one mainly uses \textit{uninterpreted} types. \textit{Uninterpreted} means that the type is an arbitrary set of values with at least one element.
+The cardinality of this set is \textit{parameterized}, meaning IVy proofs correctness for an arbitrary amount of values of an \textit{uninterpreted} type.
+The number of values of a type that are generated to construct a counter example to induction (CTI) is entirely up to IVy. No instances of these types are manually instantiated.
+This is useful for declaring types such as \textit{node}, \textit{process} or \textit{value} in a system.
 
@@
 To assign values to relations, functions or variables $:=$ is used. To assign to multiple values
-at the same time one can use an uppercase placeholder. For example setting all nodes with an incoming message to failed:
+at the same time one can use an uppercase placeholder. For example, to set all nodes with an incoming message to \textit{failed}:
 
@@
 \subsubsection{Actions}
-To define a callable action, that is callable from within another action or from the environment if it is exported
-this is the syntax.
+The syntax to define an action, that is callable from within another action, or from the environment when \textit{exported} is:
 
@@
 \subsubsection{Non deterministic choice}
-To assign to a variable, relation or function a value non-deterministically (randomly) one uses ``*''.
+To non-deterministically assign to a variable, relation or function ``*'' is used.
 This line non-deterministically chooses a number of nodes to have failed.
@@
 \subsubsection{Assume, require and ensure}
-To specify safety conditions in an action we can use require and ensure.
-They both act like a traditional assert in that it does not allow control to pass if it is false.
-They are different semantically in that require assigns blame to the actions caller, while ensure
-is a condition that the action itself must guarantee. This is especially relevant to restrict
+To specify safety conditions in an action we can use \textit{require} and \textit{ensure}.
+They both act like a traditional assert: they do not allow control to pass if they evaluate to false.
+However, they are different semantically: \textit{require} assigns blame to the actions caller, while \textit{ensure}
+is a condition that the action itself must guarantee. \textit{Require} is especially relevant to restrict
 the environment in the way it can call an action.
 
-Assume is different in that it does not allow control to pass through if the condition is not satisfied.
-The author has found this to be mostly useful in combination with non-deterministic assignments that one
-wants to restrict.
-
-The following action chooses a random node to assign to y, but will not allow control to pass if
-the chose node has failed. Therefore the actions will choose a random non failing node to send a message to x.
+\textit{Assume} is used to exclude the possibility of something happening.
+This is mostly useful in combination with non-deterministic assignments that one wants to restrict.
+The following action chooses a random node to assign to y, but will not allow control to pass if the chosen node has failed.
+Therefore the actions will choose a random non-failed node to send a message to x.
 If there is no node that hasn't failed, this action cannot terminate.
@@
 it is risky to use assume for the same reason: one might dismiss a valid counterexample to the specification
-by assuming it cannot happen.
+by assuming it cannot happen. Putting $assume\ false$ at the beginning of any action will cause all checks on it
+to be positive.
 
-\subsubsection{Invariant}
-To specify the properties we want to proof about our algorithm we encode them in inductive invariants like so:
+\subsubsection{Invariants}
+To specify the properties we want to proof about our algorithm we encode them in special assertions called \textit{invariants}:
 
@@
 \end{mdframed}
-
 These must hold at all times when an action is not executing, thus they have to be initially true and each of
-the exported actions must preserve them. However, an invariant is not guaranteed to hold when an action calls another action.
+the \textit{exported} actions must preserve them. However, an invariant is not guaranteed to hold when an action calls another action.
 
 \subsection{Decidability}
-Not all first-order formulas are decidable in Ivy. Therefore, the system defines a subset of first-order formulas as its decidable fragment.
-This not only constraints the structure of logical formuals but also functions. In Ivy only stratified functions are allowed. Meaning that if a graph was constructed
-with every type being a node and every function from a type A mapping to another type B would be represented as an edge the graph has to be cycle free. Thus there cannot be a function from
+Not all first-order formulas are decidable in IVy. Therefore, the system defines a subset of first-order formulas as its decidable fragment.
+This constraints the structure of logical formuals and functions. In IVy only stratified functions are allowed. Meaning that a graph,
+with every type being a node and every function mapping from type $A$ to type $B$ be an edge from $A$ to $B$, has to be cycle free. Thus, there cannot be a function from
 $X \rightarrow Y$ and from $Y \rightarrow X$ in the same specification.
 
-Ivy is extremely vocal about this and will let the user know if his specification does not lie in the decidable fragment.
+IVy is vocal about this and warns users about specifications that it thinks do not lie in the decidable fragment.
+However, IVy's check is conservative. It sometimes warns about specifications that are actually decidable. For this
+reason the check can be deactivated.
 
@@
 
-The algorithm that was implemented in Ivy as part of this work is called \textit{Floodset} and was taken out of Lynch's work ``Distributed Algorithms'' \cite{refNancy}.
-\textit{Floodset} is a consensus algorithm for synchronous networks that is designed to reach consensus in a network with failing processes.
-To understand \textit{Floodset}, one needs to first be made familiar with the aforementioned definition of the framework/system within which this algorithm operates
+The algorithm that was implemented in IVy as part of this work is called \textit{Floodset} and was taken out of Lynch's work ``Distributed Algorithms'' \cite{refNancy}.
+\textit{Floodset} is a consensus algorithm for synchronous networks with failing processes.
+To understand \textit{Floodset}, one needs to first understand the aforementioned \textit{framework}/\textit{system} within which this algorithm operates.
 
@@
 
-The network of processes is represented as a directed graph $G=(V,E)$. Where $n$ is $|V|$ the number of nodes in the network. Each node represents a process in the network.
+The network of processes is represented as a directed graph $G=(V,E)$. Where $n$ is $|V|$, the number of nodes in the network. Each node represents a process in the network.
 The communication paths between nodes are represented by directed edges in the graph.
-$out-nbrs_i$ is the set of nodes to which an edge from node $i$ exists.
-$in-nbrs_i$ is the set of nodes from which an edge to node $i$ exists.
+$outnbrs_i$ is the set of nodes to which an edge from node $i$ exists.
+$innbrs_i$ is the set of nodes from which an edge to node $i$ exists.
 $M$ denotes the message alphabet, meaning that elements of $M$ can be used to construct messages (vectors of elements of $M$) from nodes to other nodes.
@@
     \item $start_i$, a nonempty subset of $states_i$
-    \item $msgs_i$, a message generation function that maps $states_i X out-nbrs_i$ to elements of $M U {null}$
-    \item $trans_i$, a state transition function that maps $states_i X vectors of M index by in-nbrs_i$ to $states_i$
+    \item $msgs_i$, a message generation function that maps $states_i \times outnbrs_i$ to elements of $M \cup \{null\}$
+    \item $trans_i$, a state transition function that maps $states_i \times vectors\ of\ M$ index by $innbrs_i$ to $states_i$
 \end{itemize}
 Thus, a node is always in some state in $states_i$. Importantly, the set of states is not necessarily finite, meaning that unbounded data structures like arrays can be modeled in the system.
-Each node defines a message generation function that uses the current state to return a message for each out neighbour. Similarly each node has a state transition function that is used to determine
-from the current state and all incoming messages the state the next state the node will transition to.
+Each node defines a message generation function $msgs_i$, that uses the current state to return a message for each out neighbour.
+Similarly, each node has a state transition function $trans_i$, that is used to determine
+from the current state and all incoming messages the next state the node will transition to.
 
@@
 The system of nodes operates in steps. It begins with all nodes in an arbitrarily chosen start state from $start_i$ and all channels empty.
-The processes then all repeatedly perform the following two steps \textit{synchronously}:
+The processes then repeatedly perform the following two steps \textit{synchronously}:
 \begin{itemize}
@@
 \end{itemize}
-\textit{Synchronously} in this case means that all nodes wait for each other to finish step1 before executing step2 and vice versa.
+\textit{Synchronously} in this case means, that all nodes wait for each other to finish Step 1 before executing Step 2 and vice versa.
 The combinations of these two steps is a \textit{round}.
@@
 The system presented in Lynch's work \cite{refNancy} is capable of modelling failure in a variety of ways.
-This work makes use of stopping failure only. This kind of failure occurs when a process stops in the middle of its execution. With regards to the presented framework this
+This work makes use of stopping failure only. This kind of failure occurs when a process stops in the middle of its execution. With regards to the presented \textit{framework} this
 is modeled as a process failing before or after performing some instance of Step 1 or Step 2. Additionally, a process may also stop in the middle of performing Step 1. This allows
-the interesting situation in which a process has communicated its messages to some but not yet all processes. It is important to note that processes are not thought of performing
+the interesting situation in which a process has communicated its messages to some, but not all processes. It is important to note, that processes are not thought of performing
 the message generation operation sequentially. Instead, we say that the order in which messages are put in the channels is non-deterministic. This clarification, allows a failing process
@@
 The problem at hand is as follows: An arbitrary number of processes start with individual inputs from a value set $V$. This is modeled by each node having at least one start state
-for every possible value $v ∈ V$. Also, a default value $v_0 ∈ V$ that is equal for all nodes is defined.
+for every possible value $v \in V$. Also, a default value $v_0 \in V$ that is equal for all nodes is defined.
 Processes communicate with each other through several \textit{rounds}. The nature of their communication and its effect on their states is defined by their $msgs_i$ and $trans_i$ functions.
@@
   \item \textbf{Agreement}: No two processes decide on different values.
-  \item \textbf{Validity}: If all processes start with the same initial value $v ∈ V$, then $v$ is the only possible decision value
+  \item \textbf{Validity}: If all processes start with the same initial value $v \in V$, then $v$ is the only possible decision value
   \item \textbf{Termination}: All nonfaulty processes eventually decide.
@@
 In Step 1 of every round nodes broadcast their $W$.
-In Step 2 all $v ∈ V$ that \textit{(a)} are sent to a process and \textit{(b)} the process does not yet hold in its $W$ are added to its $W$.
+In Step 2 all $v \in V$ that \textit{(a)} are sent to a process and \textit{(b)} the process does not yet hold in its $W$ are added to its $W$.
 In the $f+1$th round
@@
 
-\noindent What follows is a proof of the correctness of the \textit{Floodset} algorithm to illustrate the principles behind it to the reader.
-
-Let $W_i(j)$ denote the value of $W$ for node $i$ in round $j$. The following Lemmas apply:\\
+\noindent What follows is a proof of the correctness of the \textit{Floodset} algorithm to illustrate the principles behind it.\\
+\\
+\noindent Let $W_i(j)$ denote the value of $W$ for node $i$ in round $j$. The following Lemmas apply:\\
 
@@
 \textbf{Proof}:\\
-\textit{Indirect}: If there exists a value $v \in V$ that is in $W_i(r)$ but not in $W_j(r)$ after round $r$ in which no process failed, that implies that some node $k$ sent a value to $i$ that it did not send to $j$, or $i$ failed to send a member of $W_i(r-1)$.
+\textit{Indirect}: If there exists a value $v \in V$ that is in $W_i(r)$ but not in $W_j(r)$ after round $r$, in which no process failed, that implies, that some node $k$ sent a value to $i$ that it did not send to $j$, or $i$ failed to send a member of $W_i(r-1)$.
 Since all nodes broadcast their $W$, this is only possible if node $k$ or $i$ failed to send at least one message.
@@
 \textbf{Proof}:\\
-Since all active nodes  have equal $W$s after $r$ rounds, there exists no $v \in V$ that is in some $W_i(r),\ i\ \mathrm{some\ active\ node}$ that is not in every other $W_j(r),\ j\ \mathrm{any\ active\ node}$.
-Since nodes will only send their $W$s, this implies that no node can send a message in some future round $r'$ that contains a value $v$ that some node does not have in its $W$, thus $W_i(r’) = W_j(r’)$, $i,\ j$ not failed.\\
+Since all active nodes have equal $W$s after $r$ rounds, there exists no $v \in V$ that is in some $W_i(r),\ \neg failed(i)$, that is not in every other $W_j(r), \neg failed(j)$.
+Since nodes will only send their $W$s, this implies that no node can send a message in some future round $r'$ that contains a value $v$ that some node does not have in its $W$, thus $W_i(r’) = W_j(r’)$, $\neg failed(i) \land \neg failed(j)$.\\
 
@@
 \textit{Indirect:} If a process failed in all of $f+1$ rounds at least on process must have failed in all of the $f+1$ rounds. This implies
-that at least $f+1$ processes failed. Only $f$ process can fail. \textit{Contradiction}.\\
+that at least $f+1$ processes failed. Only $f$ processes can fail. \textit{Contradiction}.\\
 
@@
 
-\noindent The Lemmas allow proofing the theorem:\\
+\noindent The Lemmas allow proving the theorem:\\
 \noindent \textbf{Thorem}:\\
@@
 If all processes start with the same initial value $v \in V$, then there exists no value $v', v \neq v'$, that is part of any $W$.
-Thus, from $W_i(f+1) = W_j(f+1)$, $i,\ j$ active, and $W_i(f+1) = \{v\}$ it follows
+Thus, from $W_i(f+1) = W_j(f+1)$, $\neg failed(i) \land \neg failed(j)$, and $W_i(f+1) = \{v\}$ it follows
 that every active node decides $v$.\\
@@
 
-This work now presents a way to translate Lynch's framework and the \textit{Floodset} algorithm to a specification in the Ivy language.
+This work now presents a way to translate Lynch's \textit{framework} and the \textit{Floodset} algorithm to a specification in the IVy language.
 
 \section{Specification}
-To summarize: at hand we have a parameterized system with $n$ process and $f$ number of faults.
-We want to show \textit{Floodset} is correct regardless of the parameters.
-It is trivial to verify for a small number of $n,\ f$, but the question is ``does \textit{Floodset} work for any assignment of $n,\ f, n \lt f$''.
-To answer this question we are using the theorem proofer Ivy; first encoding the algorithm in the Ivy language and then constructing inductive invariants to proof the correctness.
+To summarize: we have a parameterized system with $n$ process and $f$ number of faults. We want to show \textit{Floodset} is correct regardless of the parameters.
+It is trivial to verify for a small number of $n,\ f$, but the question is ``does \textit{Floodset} work for any assignment of $n,\ f,\ n < f$''.
+To answer this question the algorithm was encoded in the IVy language and then inductive invariants were constructed to proof correctness.
+
+The goal of this section is to walk through the specification step by step, to illustrate the paradigms relevant to automatic verification in IVy.
+The complete code, together with setup instructions of a Docker container, can be found here \cite{github}.
+
+\subsection{Modeling strategy}
+
+\subsubsection{Modularity}
+When verifying an algorithm like \textit{Floodset} in IVy, an algorithm designer would prefer to translate their original algorithm to the IVy language in as close to a one-to-one fashion as possible.
+However, this is not quite convenient for \textit{Floodset} and the \textit{framework} it operates in.
+To separate system logic and node logic one would have to rely heavily on iterations.
+E.g.
+for \textit{Step 1} the system would iterate over all nodes and let each node take care of sending it's own message.
+
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+action round = {
+  var iter := node.t.iter.create(0);
+  while ~iter.is_end
+  {
+    call send_messages(iter.val)
+    iter := iter.next;
+  }
+  . . .
+}
+action send_messages(n:node.t) = {
+  node.message(n,V) := node.w(n,V);
+  node.incomming_message(O,n) := true;
+}
+
+export round
+\end{lstlisting}
+\end{mdframed}
+
+\noindent However, the code above does not actually execute an iteration over all nodes as one would expect. Looking at an IVy trace we see what really happens:
+
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+. . .
+line 3: message := *;
+line 3: incomming_message := *;
+. . .
+\end{lstlisting}
+\end{mdframed}
+
+IVy non-deterministically assigns to all relations involved in the loop, instead of actually executing the loop.
+This is because proving correctness of loops can only be done through inductive loop invariants specified by a human.
+IVy therefore allows specifying invariants together with a loop.
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+  while ~iter.is_end
+  invariant node.message(N,V) = node.w(N,V)
+  invariant N <= iter.val -> incomming_message(O,N)
+  {
+    . . .
+  }
+\end{lstlisting}
+\end{mdframed}
+
+These invariants must hold before the loop and after every iteration.
+The following is a trace of the execution of the loop with the specified invariants.
+First, they are asserted, then the relations are set to non-deterministic values, and finally an assume is used to restrict control to relation assignments that fulfil the invariants.
+
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+assert message(N,V) = w(N,V)
+assert N:t <= t.iter.val(loc:iter) -> incomming_message(O,N)
+incomming_message := *
+message := *
+assume message(N,V) = w(N,V)
+assume N:t <= t.iter.val(loc:iter) -> incomming_message(O,N)
+\end{lstlisting}
+\end{mdframed}
+
+As a result of this architecture, the program logic has to be specified twice: once in the form of invariants and another time in the actual loop.
+Such a program therefore has two possible origins for any error: a fault in the algorithm or an inadequate invariant.
+This structure also increases the difficulty of debugging, because IVy only shows unhelpful $assert$, $assume$ and  non-determinsitic assignments ($:= *$) in its traces, rather than execution of the actual lines in the $send\_messages$ function.
+
+This approach may still be considered viable for the simple \textit{Step 1}, but for the more complex \textit{Step 2} and the failure simulation logic, the author has found that the duplication of logic and the decreased debugging efficiency makes the increased modularity achieved through iterations not worth the mental and maintenance effort that the loop invariants bring with them.
+
+For this reason the following specification of \textit{Floodset} uses a non-modular structure where single lines predominately alter the state of the system as a whole.
+
+\subsubsection{Cardinality relationships}
+There is no cardinality operator in IVy.
+In set notation we would write something like $|\{n|node(n) \land failed(n)\}|$ to count the number of failed nodes in a system.
+When programming in a standard procedural language we would use a simple loop for this task but, as has been explained in the previous section, loops are somewhat problematic in IVy.
+The author couldn't find a way to specify an inductive loop invariant for a loop counting the number of failed nodes without a cardinality operator at hand.
+This means that alternative encodings for the assumption $f<n$ and the statement $if\ rounds = f+1$ must be found.
+
+The assumption $f<n$, ``number of failed nodes is bounded to an amount smaller than the number of total nodes'', can alternatively be expressed via $\exists n\ \neg failed(N)$.
+
+To figure out how to alternatively encode $if\ rounds = f+1$ deep understanding of the algorithm is necessary.
+One must realise that $rounds = f+1$ entails that a ``clean round'' occurred (see \textit{Lemma 3}).
+Where ``clean round'' describes ``a round in which no node failed''.
+The processes are ready to decide after a clean round has occurred.
+They can decide either immediately, or an arbitrary number of rounds after the clean round.
+This can be modeled with non-determinism:
+
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+clean_round_occured := (clean_round_occured |
+                        (forall N . ~node.failed_during_round(N)));
+if clean_round_occured {
+  decision_round := *;
+};
+\end{lstlisting}
+\end{mdframed}
+
+\noindent And an invariant that describes the properties of a clean round, that contains the premise of \textit{Lemma 1}:
+
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+# lemma 1:
+invariant system.clean_round_occured ->
+            ((~node.failed(A) & ~node.failed(B)) ->
+              (node.w(A, V) = node.w(B, V)))
+\end{lstlisting}
+\end{mdframed}
+
+\noindent The (other) chosen invariants will be explained in detail later on in this paper.
+
+\subsection{Types, functions, relations}
+First the types, functions, relations and individual variables that are used throughout the program are presented:
+
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+type node
+type value
+object node = {
+  #graph/system related
+  function val(X:node) : value
+  function default(X: node) : value
+  individual def : value
+  relation failed(A:node)
+  relation failed_before_round(A:node)
+  relation failed_during_round(A:node)
+  relation has_single_value(N:node)
+
+  #state
+  relation w(SELF:node, V:value)
+  relation decision(SELF:node, VALUE:value)
+
+  #communication
+  relation message(N:node, V:value)
+  relation incomming_message(SELF:node, OTHER:node)
+  relation incomming_message_failing(SELF:node, OTHER:node)
+}
+
+object system = {
+  individual decision_round:bool
+  individual clean_round_occured:bool
+}
+\end{lstlisting}
+\end{mdframed}
+
+\noindent Their meaning will become clear as the code is presented in the following paragraphs.
+
+\noindent A round in the specification is defined as follows:
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+action round = {
+  require ~node.decision(N,V) & ~decision_round;
+  node.failed_before_round(N) := node.failed(N);
+  call step1;
+  call simulate_failure;
+  clean_round_occured := (clean_round_occured |
+                          (forall N . ~node.failed_during_round(N)));
+  if clean_round_occured {
+    decision_round := *;
+  };
+  call step2;
+}
+export round;
+\end{lstlisting}
+\end{mdframed}
+
+\noindent Line 2 makes it illegal for the environment to call \textit{round} if decisions already occurred.
+\noindent Line 3 does some bookkeeping.
+\textit{call step1} models the communication step. \textit{simulate\_failure} simulates failure before, during
+or after \textit{step1}. After \textit{decision\_round} is set, \textit{call step2} is
+executed to model the state transition of the nodes.
+
+\subsection{Step 1 - Message sending}
+The contents of a nodes message is modeled via the relation $message(N:node,V:value)$. Where
+$message(1,0) = true; message(1,1) = true;$ means node $1$ includes in its message the values $0$ and $1$.
+Since a node sends the same message to all nodes, for any given round, we don't need to include the
+recipient in the relation.
+
+The actual delivery of a message is modeled via the $incomming\_message(RECIPIENT:node,SENDER:node)$ relation.
+Where $incoming\_message(0,1) = true; incoming\_message(0,2) = true;$ means that node $0$ receives the messages from node $1$ and node $2$.
+
+These relations allow us to use IVy's assignments to easily model \textit{Step 1} in a round of \textit{Floodset}.
+Line 2 assigns for all nodes $A$ and values $V$ to $message(A,V)$ the boolean value of $node.w(A,V) \land \neg failed(A)$.
+In other words, every node that hasn't failed prepares a message with all values in its $w$.
+Line 3 states that all nodes that haven't failed send messages to all other nodes.
+
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+action step1 = {
+  node.message(A,V) := node.w(A,V) & ~node.failed(A);
+  node.incomming_message(O,A) := ~node.failed(A);
+}
+\end{lstlisting}
+\end{mdframed}
+
+\subsection{Step 2 - Message sending}
+
+Line 2 of the \textit{step2} action adds to $w$ of node $N$ all values $V$ for which a node $J$ exists that has sent a message to $N$ that contains $J$.
+The conditional statement in Line 4 corresponds to $if\ rounds = f+1$ in the original algorithm. Nodes with a single value in $w$ decide on that value, otherwise
+they decide on the default value.
+Finally, for good measure, $incomming\_message$ and $message$ are reset, but this is not necessary for the proof.
+
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+action step2 = {
+  node.w(N,V) := (node.w(N,V) |
+                 (~node.failed(N) &
+                  exists J . (node.incomming_message(N,J) &
+                               node.message(J,V))));
+
+  if decision_round {
+
+    node.has_single_value(N) :=
+      (exists J . node.w(N,J)) &
+      (forall V,O . O=V | (node.w(N,V) -> ~node.w(N,O)));
+
+    node.decision(N,K) :=
+      node.has_single_value(N) & node.w(N,K) & ~node.failed(N);
+
+    node.decision(N,node.default(N)) :=
+      (~node.has_single_value(N) | node.decision(N,node.default(N)))
+      & ~node.failed(N);
+
+  };
+
+  node.incomming_message(A,B) := false;
+  node.message(A,V) := false;
+}
+\end{lstlisting}
+\end{mdframed}
+
+\subsection{Failure simulation}
+
+The \textit{simulate\_failure} action simulates the failure model described by Lynch.
+It is called after \textit{step1} and illustrates sophisticated usage of $assume$.
+
+In line 2 we choose a random number of nodes to fail. The \textit{assume} in line 3 then excludes all
+cases where a node that has already failed was chosen to fail in this round. (A node can only fail once).
+
+Line 4 prepares for each node $M$ a non-deterministically chosen set of nodes that it sends a message to, \textit{incomming\_message\_failing}.
+Line 6 updates the \textit{incomming\_message} relation. If node $M$ failed during this round, the set of nodes
+ $M$ sends messages to is changed to the values of \textit{incomming\_message\_failing}, otherwise it remains the same.
+
+ After the nodes that \textit{failed\_during\_round} are set to \textit{failed} in line 12, line 13 assumes
+ the condition $f<n$. This excludes the cases where all nodes failed.
+
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+  action simulate_failure = {
+    node.failed_during_round(N) := *;
+    assume node.failed_during_round(N) -> ~node.failed(N);
+    node.incomming_message_failing(N,M) := *;
+
+    node.incomming_message(N,M) :=
+      (node.failed_during_round(M) &
+        node.incomming_message_failing(N,M)) |
+      (~node.failed_during_round(M) &
+        node.incomming_message(N,M));
+
+    node.failed(N) := (node.failed(N) | node.failed_during_round(N));
+    assume exists N . ~node.failed(N);
+  }
+\end{lstlisting}
+\end{mdframed}
+
+\subsection{Initialisation}
+
+Finally, we describe the initial system state in a special \textit{after init} action. When
+proving correctness, IVy requires the initialisation to entail the invariants. All relations and functions
+that are not assigned are set non-deterministically by the environment. E.g. $node.val(N)$ is
+treated as a parameter the environment sets.
+
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+after init {
+  def := *;
+  default(N) := def;
+  w(N, V) := V=node.val(N);
+  incomming_message(A, B) := false;
+  decision(N, V) := false;
+  clean_round_occured := false;
+  decision_round := false;
+  node.failed_before_round(N) := false;
+  node.failed(N) := false;
+  node.has_single_value(N) := false;
+}
+\end{lstlisting}
+\end{mdframed}
+
+\section{Verification}
+
+Now that the system has been modeled, the next step of verifying correctness with IVy is specifying inductive invariants.
+The simplest invariants to come up with are those, that describe the correctness of the algorithm at hand.
+In our case, these are \textit{agreement}, \textit{validity}, \textit{termination}.
+
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+####### Correctness Conditions #######
+
+# Agreement: No two processes decide on different values.
+invariant (~node.failed(A) & ~node.failed(B)) ->
+            (node.decision(A,V) = node.decision(B,V))
+
+# every node decides on at most one value
+invariant node.decision(N,V) -> (~exists O . node.decision(N,O) & O~=V)
+
+# Validity: If all processes start with the same initial value v,
+#           then v is the only possible decision value
+invariant (exists N,M . node.val(N) ~= node.val(M)) |
+          (system.decision_round ->
+            (~node.failed(N2) -> node.decision(N2,node.val(N2))))
+
+# Termination: All nonfaulty processes eventually decide.
+invariant system.decision_round <->
+            (forall N . (~node.failed(N) ->
+                          (exists V . node.decision(N,V))))
+\end{lstlisting}
+\end{mdframed}
+
+If one were to run $ivy\_check$ on the program without specifying any
+additional invariants, IVy finds that the \textit{agreement} invariant is not inductive,
+and produces a counter example to induction (CTI). The CTI can be drawn in a GUI.
+In the left hand section of the program we can see, that the system transitioned from state 0
+to state 1 by executing \textit{round}. In the center IVy displays state 0, meaning the state
+before \textit{round} was executed. The brown edges specify the relation $w$, the purple edges
+show $val$.
+
+\noindent\includegraphics[width=\textwidth]{gui.png}
+
+\noindent State 0 shows, that no node has their initial value, modeled by $val$,
+in its $w$. The user must now realise, that this is an unreachable state, and exclude it by coming
+up with an invariant.
+
+The invariant excluding the unreachable state says, that the initial value of a node is always in its $w$:
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+# initial value is in W
+invariant node.w(N, node.val(N))
+\end{lstlisting}
+\end{mdframed}
+
+The method for finding the remaining invariants is similar: Execute $ivy\_check$ $\rightarrow$ examine CTI $\rightarrow$ design invariant $\rightarrow$ execute $ivy\_check$.
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+# the global default is the one the nodes know
+invariant node.default(N) = node.def
+# if a value is in a process' W, it must be initial value of
+# some other process (no invented values)
+invariant node.w(N, V) -> (exists M . node.val(M)=V)
+# lemma 1:
+invariant system.clean_round_occured ->
+            (~node.failed(A) & ~node.failed(B)) ->
+              (node.w(A, V) = node.w(B, V)))
+\end{lstlisting}
+\end{mdframed}
+
+Since it is also required that the \textit{framework} behaves correctly it makes sense to include some
+correctness conditions specifically aimed at the system.
+
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+####### system correctness checks #######
+
+# f < n
+invariant exists N . ~node.failed(N)
+
+# 'after' a node has failed it cannot start working again
+invariant node.failed_before_round(N) -> node.failed(N)
+
+# 'after' a node has failed it will not send anything
+invariant node.failed_before_round(N) -> ~node.incomming_message(M, N)
+
+# If the decision round occured, a clean round must have occured
+invariant system.decision_round -> system.clean_round_occured
+
+\end{lstlisting}
+\end{mdframed}
 
@@
 
+How viable is IVy as a tool for algorithm designers to proof correctness
+of their work?  The goal, when turning to a tool like IVy, is certainty of
+correctness. However, absolute certainty is never really achieved.
+One can not trivially determine, that the specification in the IVy language is logically equivalent to the actual algorithm.
+This is compounded by the fact, that the IVy language is structurally unlike classical procedural languages. Making it challenging to achieve
+a one-to-one translation from a classical distributed algorithm in a language like C. An example of this is
+the relatively complicated quantifier-logic in assignments that update the global system state.
+A similar threat is posed by $assume$ statements. The author of this paper originally modeled the line
+
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+node.w(N,V) := (node.w(N,V) |
+               (~node.failed(N) &
+                exists J . (node.incomming_message(N,J) &
+                             node.message(J,V))));
+\end{lstlisting}
+\end{mdframed}
+
+as
+
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+node.tmp_w(A,V) := node.w(A V);
+node.w(A,V) := *;
+assume node.w(A,V) <-> (node.tmp_w(A,V) |
+                        ((node.incomming_message(A,J) &
+                          node.message(J,V))));
+\end{lstlisting}
+\end{mdframed}
+ and did not discover that the $assume$ is bogus and excludes important possibilities from being verified.
+ The error only became apparent when it was discovered that correctness could be proven in the absence of an invariant that should have been absolutely necessary for induction.
+
+Furthermore, it has already been shown, that the lack of easy iteration over types makes it difficult to achieve modularity.
+The lack of which introduces additional uncertainty about correct modelling, especially in distributed systems.
+
+However, as has been pointed out, it is not just the algorithm that must be correctly specified, but the underlying \textit{framework} is equally important to get right.
+In the presented specification it is possible to remove the \textit{simulate\_failure} action and still receive a positive check from the IVy system.
+This is unsurprising, as we do not have any mechanism in place that checks that the system correctly models the possibility for node failure.
+E.g.
+to check that the simulate failure action allows a node to fail and send a message only to a subset of other nodes one can purposefully insert an assertion that only fails when the desired behaviour is possible, e.g.
+insert
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+ensure ~(node.failed(N) &
+        (exists J . ~incoming_message(J,N) &
+          exists K . incomming_message(K,N) & J~=N & K~=J & N~=K))
+\end{lstlisting}
+\end{mdframed}
+after the \textit{simulate\_failure} action.
+If IVy does not find a counterexample to this assert, the specification is erroneous.
+In light of this, it would be beneficial to have a built in mechanism to verify the possibility of something to occur.
+In this case, it would allow one to verify, that the modeled system allows for the possibility of a node to fail such that it only sends messages to a subset of processes, as part of the final proof.
+
+Of course, a variety of bugs and exploits do not stem from a logically unsound algorithmic foundation but implementation errors.
+A wonderful example is the Heartbleed exploit in OpenSSL.
+It would have been no help to verify the simple Heartbleed protocol in IVy before implementing it in C, since the error was in an implementation detail allowing a buffer underflow.
+The point is, the implementation of the algorithm must be tested through unit testing or bounded model checkers, despite successful verification in IVy.
+
+Despite these limitations, IVy trumps similar tools in user friendliness by upping interactivity.
+IVy's counterexample to induction (CTI) viewer is remarkable, consistently showing truly minimal
+state configurations that allow the user to quickly grasp the implementation problem. The GUI also allows stepping
+through individual lines and their effect on the state. Therefore, if a user is capable of
+navigating the limitations with care, IVy is absolutely adequate for proving the logical soundness of
+an algorithmic concept.
+
 \section{Conclusion}
-This paper gave a brief introduction into Controlled Natural Languages (CNLs)
+This paper has presented the Ivy system and important Ivy language features. It classified IVy
+as part of a three step process of algorithm verification.
+It explained that its usage requires a readiness to let go of procedural programming paradigms
+for the logic oriented language. It gave an explanation of induction and inductive invariants,
+explaining what they are, how they are used by IVy to proof correctness and how they can be found.
+It presented some of the systems capabilities to verify distributed algorithms by example.
+
+It pointed out some limitations. One being that it is hard to know an algorithm was correctly specified in IVy. This is due to
+the lack of easy type iteration, which leads to complicated global assignments and a non-modular structure that is far off
+the original algorithm's. The paper suggested a mechanism to verify the possibility of a state to occur to alleviate this uncertainty.
+Another limitation was the difference between a logically sound algorithm and a correctly implemented one, and how IVy can only verify the former.
+
+It highlighted the systems strengths to be user friendliness, interactivity and a powerful logic based language that allows for
+modeling of complex system state transitions in simple one-liners.
+
+It concludes that IVy's user friendliness in presenting CTIs - so long as the user is vary of the presented limitations - enables
+finding inductive invariants for proving logical soundness of an algorithm, relatively easily.
 

### 2018-07-09 paper/paper.tex
citations

@@
 has not been effectively automated. Most of the systems currently in place require a user to annotate the
-specification with inductive invariants. IVy does not solve this problem but instead increases user friendliness.
+specification with inductive invariants. IVy does not solve this problem directly, but instead aims to increase user friendliness.
 
@@
 In light of this, the IVy system was designed to display proof-failure comprehensively to the human user, making the root cause of errors more visible.
-\cite{ivy}
+\cite[ p.1]{ivy}
 
@@
 
-This paper does not aim to be a comprehensive documentation of IVy or quantitatively compare IVy's effectiveness to other verification tools.
+This paper does not aim to be a comprehensive documentation or quantitatively compare IVy's effectiveness to other verification tools.
 
@@
 Step 2 requires the user to understand the "unusual" \cite{refLanguageDoc} - compared to a procedural programming language -  IVy language features. The user must
-find a way to use the IVy language features to encode not only the algorithm itself but also the \textit{system context} or \textit{framework} in which the algorithm was designed to execute.
+find a way to use the IVy language to encode not only the algorithm itself but also the \textit{system context} or \textit{framework} in which the algorithm was designed to execute.
 
@@
 This process is fully creative and will most likely not allow a mindless one-to-one translation from a procedural algorithm.
-The user also has to tell IVy which methods (\textit{actions}) the environment can execute, e.g. \textit{connect\_to\_server(node)}. These actions are called \textit{exported}.
-These are the actions and end-user of the algorithm would call. It is our desire to proof correctness of these actions.
+The user also has to tell IVy which methods (\textit{actions}) the environment can execute, e.g. \textit{connect\_to\_server(node)}.
+These actions are called \textit{exported}, and constitute the actions an end-user of the algorithm would call. It is our desire to proof correctness of these actions.
 \subsubsection{Step 3 - verification}
-The verification happens in step 3. The goal is to specify inductive invariants that describe the properties to be proven about the algorithm. \textit{Inductive} is an important
-keyword here. An invariant is \textit{inductive} iff the following three properties are satisfied:
+The verification happens in step 3. The goal is to specify an inductive invariant that describes the properties to be proven about the algorithm. \textit{Inductive} is an important
+keyword here. An invariant is \textit{inductive} iff the following three properties are satisfied \cite{invariants}:
 \begin{enumerate}
   \item initiation: the invariant is true in all legal initial states
-  \item safety: for any program state that satisfies the invariant - regardless of whether this state is reachable - no \textit{exported} action can cause an invariant failure starting in that state
+  \item safety: for any program state that satisfies the invariant - regardless of whether this state is reachable - no \textit{exported} action can cause an assertion failure starting in that state
   \item consecution: for any program state that satisfies the invariant - regardless of whether this state is reachable - after executing any \textit{exported} action, the invariant remains satisfied
@@
 IVy does not consider the reachability of a system state.  As long as the state is in compliance with the invariants it is assumed to be reachable.
-The author of this paper needed some time to understand why IVy generates counterexamples that are clearly not reachable, e.g. negative values for counters that are set to zero at initialization.
-Every counter needs an invariant, e.g. $invariant\ counter \geq 0$.
+Thus, when IVy generates counterexamples that are clearly not reachable, e.g. negative values for counters that are set to zero at initialization,
+the solution is to specify an invariant that excludes these state, e.g. $invariant\ counter \geq 0$.
 
-The name "IVy" stems from "interactively verifying", which highlights the key differentiator of IVy to Coq or Dafny.
+When we say that the specified invariants need to be \textit{inductive} we mean the conjunction of the set of invariants specified throughout the entire program,
+not each invariant individually.
+
+The name "IVy" stems from "interactively verifying", which highlights the key differentiator of IVy to Coq or Dafny.\cite{ivy}
 IVy actively supports the user in the search for an inductive invariant by graphically displaying a concrete minimal counterexample to induction (CTI).
@@
 invariant system.rounds >= 0
+invariant system.rounds < 100
 \end{lstlisting}
 \end{mdframed}
-These must hold at all times when an action is not executing, thus they have to be initially true and each of
+The conjunction of these must hold at all times when an action is not executing, thus they have to be initially true and each of
 the \textit{exported} actions must preserve them. However, an invariant is not guaranteed to hold when an action calls another action.
@@
 However, IVy's check is conservative. It sometimes warns about specifications that are actually decidable. For this
-reason the check can be deactivated.
+reason IVy allows the check to be deactivated. \cite{decid}
 
@@
 
-\noindent And an invariant that describes the properties of a clean round, that contains the premise of \textit{Lemma 1}:
+\noindent And an invariant that describes the properties of a clean round, which effectively is \textit{Lemma 1}:
 
@@
 the relatively complicated quantifier-logic in assignments that update the global system state.
-A similar threat is posed by $assume$ statements. The author of this paper originally modeled the line
+
+Another threat is posed by $assume$ statements. The author of this paper originally modeled the line
 
@@
 
+It has become apparent, that successful verification with IVy is only possible with deep knowledge about how the algorithm functions. It is not a ``translate and go'' solution.
+Evidence of this include the complex invariant that requires knowledge about \textit{Lemma 1} or the alternative encoding for $f<n$ and the decision condition $if\ rounds = f+1$ that had to be discovered.
+
 Of course, a variety of bugs and exploits do not stem from a logically unsound algorithmic foundation but implementation errors.
@@
 
-Despite these limitations, IVy trumps similar tools in user friendliness by upping interactivity.
+Despite these limitations, IVy trumps similar tools in user friendliness by upping interactivity. \cite{ivy}
 IVy's counterexample to induction (CTI) viewer is remarkable, consistently showing truly minimal
-state configurations that allow the user to quickly grasp the implementation problem. The GUI also allows stepping
-through individual lines and their effect on the state. Therefore, if a user is capable of
-navigating the limitations with care, IVy is absolutely adequate for proving the logical soundness of
-an algorithmic concept.
+state configurations that allow the user to quickly grasp the problem. The GUI also allows stepping
+through individual lines and their effect on the state.
+
+Therefore, if a user is capable of navigating the limitations with care, IVy is absolutely adequate for proving the logical soundness of
+an algorithm.
 
@@
 It explained that its usage requires a readiness to let go of procedural programming paradigms
-for the logic oriented language. It gave an explanation of induction and inductive invariants,
+in favor of the logic oriented language. It gave an explanation of induction and inductive invariants,
 explaining what they are, how they are used by IVy to proof correctness and how they can be found.
@@
 the lack of easy type iteration, which leads to complicated global assignments and a non-modular structure that is far off
-the original algorithm's. The paper suggested a mechanism to verify the possibility of a state to occur to alleviate this uncertainty.
-Another limitation was the difference between a logically sound algorithm and a correctly implemented one, and how IVy can only verify the former.
+the original algorithm's.
+The paper suggested a mechanism to verify the possibility of a state to occur, to alleviate this uncertainty.
+Another limitation that was found, is the difference between a logically sound algorithm and a correctly implemented one, and how IVy can only verify the former.
+It was stressed that successful verification of an algorithm that is not thoroughly understood by the user is unlikely.
 
@@
 It concludes that IVy's user friendliness in presenting CTIs - so long as the user is vary of the presented limitations - enables
-finding inductive invariants for proving logical soundness of an algorithm, relatively easily.
+finding inductive invariants for proving logical soundness of an algorithm relatively easily.
 
@@
 \begin{thebibliography}{99}
-
-    \bibitem{refLanguageDoc} http://microsoft.github.io/ivy/language.html
-    \bibitem{refNancy} http://microsoft.github.io/ivy/language.html
+    \bibitem{ivy} O. Padon, K. McMillan, A. Panda, M. Safiv, S. Shoham, ``Ivy: Safety Verification by Interactive Generalization'', Available \textit{https://www.cs.tau.ac.il/~odedp/ivy.pdf}
+    \bibitem{refLanguageDoc} ``The IVy language'',  \textit{http://microsoft.github.io/ivy/language.html} [Accessed: 5.06.18], 2017
+    \bibitem{invariants} ``Invariants'',  \textit{http://microsoft.github.io/ivy/examples/client\_server\_example.html} [Accessed: 5.06.18], 2017
+    \bibitem{decid} ``Decidability'',  \textit{http://microsoft.github.io/ivy/decidability.html} [Accessed: 5.06.18], 2017
+    \bibitem{refNancy} N. Lynch, ``Distributed algorithms'', 1996, San Francisco: Morgan Kaufmann Publishers
+    \bibitem{github} C. Claus, ``BA'', \textit{https://github.com/calvinclaus/ba}, 2018
 \end{thebibliography}

### 2018-08-05 paper/paper.tex
explaining in a paragraph why it is useful to verify distributed algorithms ..as they are designed and proved by hand. Then, in a second paragraph a short explanation about the difference of automatic and deductive verification, and how it is unfeasible to apply automatic verification to distributed algorithms (due to them being parameterized, and parameterized verification is undecidable in general [1]).
Finally, say how Ivy is tailored to dealing with
these kinds of algorithms, and is a tool that allows the user to provide
invariants and check them automatically.

Also, an outline of the res

@@
 
-\lettrine[nindent=0em,lines=3]{I} t's a fascinating idea: taking any (distributed) algorithm and feeding it to
-some powerful tool that evaluates it's correctness automatically. However, the problem of verifying infinite state systems
-has not been effectively automated. Most of the systems currently in place require a user to annotate the
-specification with inductive invariants. IVy does not solve this problem directly, but instead aims to increase user friendliness.
+\lettrine[nindent=0em,lines=3]{D}istributed algorithms is a class of algorithms that is designed to run on a set of distributed processors.
+They solve important problems in distributed systems such as leader election, resource allocation and \textbf{consensus}.
+Consensus algorithms aim to create agreement among a set of processors.
+For example, an airplane may posses a redundant set of sensors measuring the same quantity.
+They must finally agree on a value that is as accurate as possible, while being able to handle failure of a subset of sensors.
+To establish agreement between sensors a distributed consensus algorithm may be used.
+Currently these algorithms are proven by hand.
+In light of the relevance of distributed algorithms, it is useful to automatise the process of proving their correctness.
 
-IVy's underlying hypothesis is that ``automated methods are difficult to apply in practice not primarily because they are unreliable, but because they are opaque''; meaning that they fail in ways inexplicable to the user.
+  \textit{Automatic verification} takes an algorithm and exhaustively checks if the algorithm satisfies some specification.
+While this approach is conveniently fully automatic it is not applicable to infinite state systems as their state-space can, by definition, not be exhaustively searched.
+It is therefore not generally applicable to \textit{parameterized} distributed (consensus) algorithms, which take the number of processors communicating as a parameter.\cite{limits}
+
+  On the other hand, \textit{deductive verification} takes an algorithm and some invariant.
+It proves the invariant is inductive with respect to the algorithm.
+If the invariant implies the correctness of the algorithm, the algorithm has been proven correct.
+This process does not exhaustively search all possible states and is therefore applicable to distributed (consensus) algorithms, however it relies on the user to annotate invariants.
+
+  Ivy is tailored to simplify deductive verification of parameterized distributed algorithms.
+There are a variety of systems currently implementing the deductive verification process.
+IVy makes use of the same underlaying principle, but it aims to increase user friendliness of the verification process.
+IVy's hypothesis is that ``automated methods are difficult to apply in practice not primarily because they are unreliable, but because they are opaque''; meaning that they fail in ways inexplicable to the user.
 In light of this, the IVy system was designed to display proof-failure comprehensively to the human user, making the root cause of errors more visible.
@@
 It will explain the paradigms underlying the automatic verification tool, its language and the context in which it may be employed.
-
 \textit{Floodset}, a simple distributed algorithm, will be explained and used as an example to illustrate the strengths and limitations of the IVy system in verifying algorithms operating in a network.
+First \textit{Floodset} will be manually proven and then deductively verified.
 The goal of this work is to give readers a foundational understanding of IVy and a head start in verifying their own distributed algorithms.
@@
 \begin{thebibliography}{99}
+    \bibitem{limits} K. Apt, D. Kozen ``Limits of Automatic Verification of Finite-State Concurrent Systems'', 1986, Available \textit{http://www.cs.cornell.edu/~kozen/Papers/LimitsForAutomatic.pdf}
     \bibitem{ivy} O. Padon, K. McMillan, A. Panda, M. Safiv, S. Shoham, ``Ivy: Safety Verification by Interactive Generalization'', Available \textit{https://www.cs.tau.ac.il/~odedp/ivy.pdf}

### 2018-08-06 paper/paper.tex
secion || 
 introductory paragraph: a few sentences about how IVy   	works could be useful at the beginning of this section. Consider moving the   	last paragraph from section (II.i.3) to this introductory paragraph  	  	- page 2, section (II.i): as we are not interested in algorithm design, but   	only in the verification of the designed algorithm, I suggest that the step 1   	of the workflow is omitted, and that the workflow is called "algorithm   	verification workflow"  	  	- page 2, section (II.i.2): the IVy language is a specification language, its   	code is not executed, so its feature

@@
 \section{IVy}
-IVy is a system that is designed to verify correctness of infinite-sate systems.
-An algorithm or protocol is \textit{correct} or \textit{safe} if it cannot be called in any way that causes some pre-defined properties (assertions, invariants) not to be satisfied.
-IVy aims to assist protocol architects, algorithm authors, etc. in automatically proving the correctness of their algorithm or protocol.
+IVy is a system designed to verify correctness of infinite-sate systems using the aforementioned deductive approach.
+It takes as input a specification of an algorithm operating on system-state  and a set of invariants.
+It proofs the inductiveness of the invariants in respect to the algorithm, or displays a counterexample to induction.
+If the set of invariants entails the desired properties of the algorithm, and Ivy proved induction, then it was proven that the algorithm is \textit{correct}.
+
+In Ivy, the entity ``calling'' the algorithm in different state configurations is called the ``environment''.
+The environment conceptually represents the final users of the algorithm.
+An algorithm or protocol is \textit{correct} or \textit{safe} if it cannot be called by the environment in a state configuration that causes some pre-defined properties (assertions, invariants) not to be satisfied.
+IVy supports verification of systems that have multiple methods (\textit{actions}) exported to the environment.
+For example to verify servers that specify connection and disconnection algorithms.
+This can be thought of as proofing inductiveness of two independent algorithms with respect to the same invariants and system-state.
 
@@
 
-A common protocol or algorithm creation workflow, which incorporates IVy for verification, is a three step process:
+A common algorithm verification workflow, which incorporates IVy for verification, is a two step process:
 
 \begin{enumerate}
-  \item algorithm or protocol design
-  \item encoding/specification in IVy language
+  \item encoding/specification of algorithm in IVy language
   \item correctness and safety verification
 \end{enumerate}
-%\begin{itemize}
-%  \item purely creative task
-%  \item IVy offers no support here
-%\end{itemize}
-%\begin{itemize}
-%  \item a one to one translation is most likely not possible
-%  \item user must be ready for the paradigm change when writing IVy language as opposed to a procedural programming language
-%\end{itemize}
-%\begin{itemize}
-%  \item specify correctness and safety conditions as invariants
-%  \item make invariants inductive
-%\end{itemize}
 
-\subsubsection{Step 1 - algorithm design}
-Step 1 is executed entirely independently of IVy. The IVy language or the IVy system do not provide any support in actually coming up with an algorithmic solution to a problem.
-The step is mentioned here to show the context the IVy system can be used in: to answer the question "is an algorithm correct"?
+\subsubsection{Step 1 - encoding in IVy language}
+Step 1 requires the user to understand the Ivy language.
+This language is a specification language tailored to describe state changes in a system.
+Ivy language code is inherently different from procedural languages, as it serves a different purpose and its logic is never executed.
+The Ivy documentation itself calls the language features "unusual" \cite{refLanguageDoc} so a user may need considerable effort to understand how to best use its tools.
 
-\subsubsection{Step 2 - encoding in IVy language}
-Step 2 requires the user to understand the "unusual" \cite{refLanguageDoc} - compared to a procedural programming language -  IVy language features. The user must
-find a way to use the IVy language to encode not only the algorithm itself but also the \textit{system context} or \textit{framework} in which the algorithm was designed to execute.
+Using their understanding of the Ivy language users have to encode their algorithm in it.
+However, distributed algorithms rarely exist in a vacuum.
+Algorithms such as \textit{Floodset} are defined within a \textit{system context} or \textit{framework} \cite{refNancy}, and make no sense without first defining this \textit{framework}.
+Thus users have to encode their algorithm together with the \textit{framework} in which the algorithm was designed to execute.
+This may include things such as network topology, the amount of processors that may fail, how and when processors fail, details about message sending, etc.
 
-For example, to reason about distributed consensus algorithms Lynch has defined a \textit{framework} in which distributed algorithms operate.
-This is explained in detail later on in this work. Part of this \textit{framework} describes that the network is defined as a graph, where edges can hold messages.
-The system executes two distinct steps: one in which messages can be sent and another in which the states of nodes are altered.
-The point is, both the algorithm and the \textit{framework} within which it operates have to be flawlessly described through the IVy language in this step.
+The encoding process therefore not only requires a clear understanding of how the algorithm works, but also of all assumptions about the context in which it executes.
+This process is therefore fully creative, requires deep understanding of the system as a whole, and will most likely not allow a mindless one-to-one translation from a procedural algorithm.
 
-This process is fully creative and will most likely not allow a mindless one-to-one translation from a procedural algorithm.
-The user also has to tell IVy which methods (\textit{actions}) the environment can execute, e.g. \textit{connect\_to\_server(node)}.
-These actions are called \textit{exported}, and constitute the actions an end-user of the algorithm would call. It is our desire to proof correctness of these actions.
-\subsubsection{Step 3 - verification}
-The verification happens in step 3. The goal is to specify an inductive invariant that describes the properties to be proven about the algorithm. \textit{Inductive} is an important
+\subsubsection{Step 2 - verification}
+The verification happens in step 2. The goal is to specify an inductive invariant that describes the properties to be proven about the algorithm. \textit{Inductive} is an important
 keyword here. An invariant is \textit{inductive} iff the following three properties are satisfied \cite{invariants}:
@@
 IVy does not consider the reachability of a system state.  As long as the state is in compliance with the invariants it is assumed to be reachable.
-Thus, when IVy generates counterexamples that are clearly not reachable, e.g. negative values for counters that are set to zero at initialization,
-the solution is to specify an invariant that excludes these state, e.g. $invariant\ counter \geq 0$.
+Thus, when IVy generates counterexamples that are clearly not reachable, e.g.
+negative values for counters that are set to zero at initialization, the solution is to specify an invariant that excludes these state, e.g.
+$invariant\ counter \geq 0$.
 
-When we say that the specified invariants need to be \textit{inductive} we mean the conjunction of the set of invariants specified throughout the entire program,
-not each invariant individually.
+When we say that the specified invariants need to be \textit{inductive} we mean the conjunction of the set of invariants specified throughout the entire program, not each invariant individually.
 
@@
 IVy actively supports the user in the search for an inductive invariant by graphically displaying a concrete minimal counterexample to induction (CTI).
-The user can then generalize from the counterexample to strengthen the invariants. This counterexample $\rightarrow$ strengthened invariant $\rightarrow$ counterexample process continues
+The user can then generalize from the counterexample to strengthen the invariants.
+This \textit{counterexample} $\rightarrow$ \textit{strengthened invariant} $\rightarrow$ \textit{counterexample} process continues
 until an inductive invariant is found and the correctness of the algorithm is proven.

### 2018-08-06 paper/paper.tex
- page 4, section (II.ii.2): explain the difference between upper and lower 	case placeholders in more detail

@@
 
-\subsubsection{Assignments}
-
-To assign values to relations, functions or variables $:=$ is used. To assign to multiple values
-at the same time one can use an uppercase placeholder. For example, to set all nodes with an incoming message to \textit{failed}:
-
-\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
-\begin{lstlisting}
-failed(X) := ~incoming_message(X);
-\end{lstlisting}
-\end{mdframed}
-
 \subsubsection{Actions}
@@
 
-action fail_node_and_send(x:node) = {
-  fail(x) := true;
-  call send;
+action exported_action(x:node) = {
+  call internal_action;
 }
-export fail_node_and_send
+export exported_action
 
-action send = {
-  sent(X) := true;
+action  internal_action = {
+  . . .
 }
@@
 \end{mdframed}
+
+\subsubsection{Assignments}
+
+To assign values to relations, functions or variables $:=$ is used.
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+action our_action(x:node) = {
+  failed(x) := true;
+}
+\end{lstlisting}
+\end{mdframed}
+Lower case tokens refer to an individual value.
+In the code above \textit{x} refers to the specific instance of type \textit{node} that has been passed to \textit{our\_action}.
+The value of the relation \textit{failed(x)} is changed to \textit{true} only for that instance of \textit{node}.
+
+In Ivy one often wants to specify system-wide state-changes.
+For this reason Ivy allows iterator-free assignments involving all instances of a type.
+To make such an assignment one can use placeholders that are denoted by uppercase tokens.
+For example, to set all nodes without an incoming message to \textit{failed}:
+
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+failed(X:node) := ~incoming_message(X:node);
+\end{lstlisting}
+\end{mdframed}
+
+\noindent This is conceptually the same as the following pseudo-code:
+
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+# Pseudocode only:
+foreach x of type node {
+  failed(x) := ~incoming_message(x);
+}
+\end{lstlisting}
+\end{mdframed}
+
+The language only supports assignments where each placeholder that appears on the right hand side of an assignment is also on the left hand side.
 
@@
 This is mostly useful in combination with non-deterministic assignments that one wants to restrict.
-The following action chooses a random node to assign to y, but will not allow control to pass if the chosen node has failed.
+The following action chooses a random node to assign to \textit{y}, but will not allow control to pass if the chosen node has failed.
 Therefore the actions will choose a random non-failed node to send a message to x.
@@
 
-Assume simplifies modelling in some situations, as it can easily eliminate impossible behaviour. However,
-it is risky to use assume for the same reason: one might dismiss a valid counterexample to the specification
-by assuming it cannot happen. Putting $assume\ false$ at the beginning of any action will cause all checks on it
-to be positive.
+\textit{Assume} simplifies modelling in some situations, as it can easily eliminate impossible behaviour. However,
+it is risky to use \textit{assume} for the same reason: one might dismiss a valid counterexample to the specification
+by assuming it cannot happen. To give an extreme example, putting $assume\ false$ at the beginning of an exported action will cause all of Ivy's checks for induction to come out positive.
 

### 2018-08-06 paper/paper.tex
- page 5, section (II.iii): first order logic is undecidable in general, but 	still many interesting properties can be expressed in a decidable fragment. 	Briefly mention which decidable fragments are captured by IVy. It is not clear 	to me what the last paragraph on this page states.

@@
 \subsection{Decidability}
-Not all first-order formulas are decidable in IVy. Therefore, the system defines a subset of first-order formulas as its decidable fragment.
-This constraints the structure of logical formuals and functions. In IVy only stratified functions are allowed. Meaning that a graph,
-with every type being a node and every function mapping from type $A$ to type $B$ be an edge from $A$ to $B$, has to be cycle free. Thus, there cannot be a function from
-$X \rightarrow Y$ and from $Y \rightarrow X$ in the same specification.
-
-IVy is vocal about this and warns users about specifications that it thinks do not lie in the decidable fragment.
-However, IVy's check is conservative. It sometimes warns about specifications that are actually decidable. For this
-reason IVy allows the check to be deactivated. \cite{decid}
+First order logic is undecidable in general, but still many interesting properties can be expressed in a subset of first oder logic called the ``decidable fragment''.
+Ivy's documentation defines this subset of first-order formulas.
+\cite{decid}
+
+For example, Ivy only allows stratified functions.
+Meaning that a graph, with every type being a node and every function mapping from type $A$ to type $B$ be an edge from $A$ to $B$, has to be cycle free.
+Thus, there cannot be a function from $X \rightarrow Y$ and from $Y \rightarrow X$ in the same specification.
+
+A detailed discussion about decidability lies outside of the scope of this work and is well documented in \cite{decide}.
+
+While a complete understanding of decidability in Ivy is advantageous, to use Ivy successfully it is not strictly necessary.
+This is because Ivy assists the user and tries to automatically determine if formulas lie outside of the decidable fragment.
+It warns users about specifications that it estimates are undecidable.
+It must be noted that IVy's check is conservative as ``fully unfolding all of the definitions could be exponential''.
+Ivy may report specifications as undecidable which are in fact decidable.
+For this reason IVy allows the check to be deactivated.
+\cite{decid}
 

### 2018-08-06 paper/paper.tex
- page 6, section (III.i): mention that the presented framework is taken from 	Lynch's book.

@@
 \textit{Floodset} is a consensus algorithm for synchronous networks with failing processes.
-To understand \textit{Floodset}, one needs to first understand the aforementioned \textit{framework}/\textit{system} within which this algorithm operates.
+To understand \textit{Floodset}, one needs to first understand the aforementioned \textit{framework}/\textit{system} within which this algorithm operates, which is also defined in \cite{refNancy}.
 

### 2018-08-06 paper/paper.tex
- page 6, section (III.i): mention that the failing processes cannot restart
- page 7, section (III.iii): the algorithm runs on n processes, in the 	presence of f faults, with n > f, both n and f are parameters, and natural 	numbers. The first paragraph of this section is a mix of explanation of the 	algorithm and of the modeling. I would split the explanation of the algorithm 	and the details about modeling in two paragraphs. It is not clear to me what 	"complete knowledge over the network" means

@@
 the message generation operation sequentially. Instead, we say that the order in which messages are put in the channels is non-deterministic. This clarification, allows a failing process
-to send its messages to an arbitrary subset of its outgoing neighbours only.
+to send its messages to an arbitrary subset of its outgoing neighbours only. A process that has failed cannot restart at a later point in time.
 
 \subsection{Floodset}
-The problem at hand is as follows: An arbitrary number of processes start with individual inputs from a value set $V$. This is modeled by each node having at least one start state
-for every possible value $v \in V$. Also, a default value $v_0 \in V$ that is equal for all nodes is defined.
-Processes communicate with each other through several \textit{rounds}. The nature of their communication and its effect on their states is defined by their $msgs_i$ and $trans_i$ functions.
-Nodes may fail according to the \textit{stopping failure} model described above. The total number of failing processes is bounded by a constant $f$. The entire network knows the value of $f$ and
-can therefore utilise it in their state transition function.
-All processes that did not fail must eventually decide on a single value from the value set $V$.
-\textit{Deciding} is modeled by nodes setting a special \textit{decision} state component to a value in $V$.
-Their decision is subject to the following correctness conditions:
+The problem \textit{Floodset} solves is as follows: An arbitrary number of nodes (or processes) $n > 0$, in a complete, undirected network with fully reliable channels, start with individual inputs from a value set $V$ and a default value $v_0 \in V$ that is equal for all nodes.
+Processes should communicate with each other through several \textit{rounds} to eventually decide on a single value $v_d \in V$.
+Nodes may fail according to the \textit{stopping failure} model.
+The total number of failing processes is bounded by a constant $f < n$.
+The entire network knows the value of $f$.
+The network should satisfy the following conditions:
+
 \begin{itemize}
@@
 \end{itemize}
-We assume the processes are in a complete, undirected graph. This means that every node has complete knowledge over the network. Channels are fully reliable - every message gets delivered.
-
-In the \textit{Floodset} algorithm processes hold a set $W$ as one of their state components.
-This set holds the subset of values of $V$ that the process has seen.
+
+We refer to these condition as ``correctness conditions'', as we say the algorithm is ``correct'' if we can proof it satisfies these conditions.
+\textit{Floodset} solves the described problem by defining $msgs$ and $trans$ functions for nodes in the network such that the conditions are satisfied.
+
+\textbf{Informal explanation of \textit{Floodset}}: Processes hold a set $W$ as one of their state components which holds the subset of values of $V$ that the process has seen.
 Every node's initial $W$ holds only its initial value.
@@
 In Step 2 all $v \in V$ that \textit{(a)} are sent to a process and \textit{(b)} the process does not yet hold in its $W$ are added to its $W$.
-In the $f+1$th round
-a node decides: If $W$ is a singleton set then it decides on the unique element of $W$, otherwise it decides on the default value $v_0$.
-
+In the $f+1$th round a node decides: If $W$ is a singleton set then it decides on the unique element of $W$, otherwise it decides on the default value $v_0$.
+\textit{Deciding} is modeled by nodes setting a special \textit{decision} state component to a value in $V$.
+
+\textbf{Formal definition of \textit{Floodset}}:
 \begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]

### 2018-08-06 paper/paper.tex
- page 8, section (III.iii): explicitly state that this algorithm has been 	proven correct in Lynch's book, and that the proofs are shortly presented 	here as well

@@
 
-\noindent What follows is a proof of the correctness of the \textit{Floodset} algorithm to illustrate the principles behind it.\\
+\noindent What follows is a proof of the correctness of the \textit{Floodset} algorithm to illustrate the principles behind it.
+It has already been proven correct in \cite{refNancy}.
+This work presents a slightly varied approach to the proofs of \cite{refNancy} founded on the same general proof ideas.\\
 \\

### 2018-08-06 paper/paper.tex
- page 9, section (IV.i): I would add an introductory paragraph for the 	section "Modeling strategy", and state that the remainder of the section 	contains some challenges that came across while using IVy, and how they were 	overcome

@@
 
-The goal of this section is to walk through the specification step by step, to illustrate the paradigms relevant to automatic verification in IVy.
+\subsection{Modeling strategy}
+
+This section walks through the specification step by step.
+It illustrates the paradigms relevant to verification in IVy and presents the challenges that the author came across while specifying in IVy, and how they were overcome.
 The complete code, together with setup instructions of a Docker container, can be found here \cite{github}.
-
-\subsection{Modeling strategy}
 

### 2018-08-06 paper/paper.tex
- page 10, section (IV.i.2): instead of using the code "if round ...", maybe 	refer to it as the "decision condition", or "decision guard". Also, explain 	the why the formula \exists N. ~failed(N) captures f < n more intuitively

@@
 In set notation we would write something like $|\{n|node(n) \land failed(n)\}|$ to count the number of failed nodes in a system.
-When programming in a standard procedural language we would use a simple loop for this task but, as has been explained in the previous section, loops are somewhat problematic in IVy.
+When programming in a procedural language one would use a loop for this task but, as has been explained in the previous section, loops are somewhat problematic in IVy.
 The author couldn't find a way to specify an inductive loop invariant for a loop counting the number of failed nodes without a cardinality operator at hand.
-This means that alternative encodings for the assumption $f<n$ and the statement $if\ rounds = f+1$ must be found.
+This means that alternative encodings for the assumption $f<n$ and the \textit{decision condition} ($if\ rounds = f+1$) must be found.
 
 The assumption $f<n$, ``number of failed nodes is bounded to an amount smaller than the number of total nodes'', can alternatively be expressed via $\exists n\ \neg failed(N)$.
-
-To figure out how to alternatively encode $if\ rounds = f+1$ deep understanding of the algorithm is necessary.
+This is because the existence of a functioning node implies that the number of failed nodes $f$ is smaller than $n$.
+
+To figure out how to alternatively encode the \textit{decision condition} deep understanding of the algorithm is necessary.
 One must realise that $rounds = f+1$ entails that a ``clean round'' occurred (see \textit{Lemma 3}).

### 2018-08-06 paper/paper.tex
- page 11, section (IV.ii): give more intuition about the encoding. Maybe 	explain in words the meaning of one function/relation of each 	graph/state/communication parts of the node

@@
 \subsection{Types, functions, relations}
-First the types, functions, relations and individual variables that are used throughout the program are presented:
+First the types, functions, relations and individual variables that are used throughout the program are presented.
+They are annotated with comments explaining their purpose.
+Their meaning will become more clear as the code is presented in the following paragraphs.
 
@@
 object node = {
-  #graph/system related
+
+#graph/system
+  # stores the initial value of each X
   function val(X:node) : value
+
+  # stores the default value of the network
+  individual def : value
+
+  # stores the default value that each node stores independently
   function default(X: node) : value
-  individual def : value
+
+  # failed(a) is true if node a failed
   relation failed(A:node)
+
+#state
+  # w(n, v) denotes whether n's w contains value v
+  relation w(SELF:node, V:value)
+
+  # decision(n, v) denotes whether n has decided on v
+  relation decision(SELF:node, VALUE:value)
+
+#communication
+  # message(n,v) denotes whether v is broadcast by n
+  relation message(N:node, V:value)
+
+  # incomming_message(n,o) denotes whether node n received a message
+  # from node o
+  relation incomming_message(SELF:node, OTHER:node)
+
+# internal helpers
+  # true if a node had already failed before this round
   relation failed_before_round(A:node)
+
+  # true if a node failed during this round
   relation failed_during_round(A:node)
+
+  # true if a node has a single value in w
   relation has_single_value(N:node)
 
-  #state
-  relation w(SELF:node, V:value)
-  relation decision(SELF:node, VALUE:value)
-
-  #communication
-  relation message(N:node, V:value)
-  relation incomming_message(SELF:node, OTHER:node)
+  # helper to simulate non-determinstic message sending
   relation incomming_message_failing(SELF:node, OTHER:node)
+
 }
@@
 object system = {
+  # true if current round is decision round
   individual decision_round:bool
+
+  # true if a clean round occured in the past, or the current round
   individual clean_round_occured:bool
@@
 \end{mdframed}
-
-\noindent Their meaning will become clear as the code is presented in the following paragraphs.
 

### 2018-08-11 paper/paper.tex
- page 13: both section (IV.iii) and (IV.iv) are called "message sending", 	probably the second one should be called "state transition"
- page 14, section (IV.v): in the text, you are referring to commands by line 	number, but the lines where the actual code is are not numbered.  Also, this 	section lacks some intuition. Try to explain what subtleties there are when 	modeling the faults, and then focus on how it is modeled using the IVy code

@@
 \lstset{
-  xleftmargin=0em,
+  xleftmargin=1em,
+  numbers=left,
+  firstnumber=1,
+  numberfirstline=true
 }
@@
 
+\subsection{Rounds}
+
 \noindent A round in the specification is defined as follows:
@@
 
+\noindent \textit{round} is the only exported action. It specifies the basic structure of Lynch's \textit{framework}.
 \noindent Line 2 makes it illegal for the environment to call \textit{round} if decisions already occurred.
@@
 \textit{call step1} models the communication step. \textit{simulate\_failure} simulates failure before, during
-or after \textit{step1}. After \textit{decision\_round} is set, \textit{call step2} is
-executed to model the state transition of the nodes.
+and after \textit{step1}. After \textit{decision\_round} is set, \textit{call step2} is executed to model the state transition of the nodes.
 
 \subsection{Step 1 - Message sending}
-The contents of a nodes message is modeled via the relation $message(N:node,V:value)$. Where
+The contents of a node's message is modeled via the relation $message(N:node,V:value)$. Where
 $message(1,0) = true; message(1,1) = true;$ means node $1$ includes in its message the values $0$ and $1$.
@@
 
-\subsection{Step 2 - Message sending}
-
-Line 2 of the \textit{step2} action adds to $w$ of node $N$ all values $V$ for which a node $J$ exists that has sent a message to $N$ that contains $J$.
-The conditional statement in Line 4 corresponds to $if\ rounds = f+1$ in the original algorithm. Nodes with a single value in $w$ decide on that value, otherwise
+\subsection{Step 2 - State transition}
+
+Line 2 in the \textit{step2} action adds to $w$ of node $N$ all values $V$ for which a node $J$ exists that has sent a message to $N$ that contains $J$.
+The conditional statement in Line 4 corresponds to the \textit{decision condition} ($if\ rounds = f+1$) in the original algorithm. Nodes with a single value in $w$ decide on that value, otherwise
 they decide on the default value.
@@
 
-The \textit{simulate\_failure} action simulates the failure model described by Lynch.
-It is called after \textit{step1} and illustrates sophisticated usage of $assume$.
-
-In line 2 we choose a random number of nodes to fail. The \textit{assume} in line 3 then excludes all
-cases where a node that has already failed was chosen to fail in this round. (A node can only fail once).
-
-Line 4 prepares for each node $M$ a non-deterministically chosen set of nodes that it sends a message to, \textit{incomming\_message\_failing}.
-Line 6 updates the \textit{incomming\_message} relation. If node $M$ failed during this round, the set of nodes
- $M$ sends messages to is changed to the values of \textit{incomming\_message\_failing}, otherwise it remains the same.
-
- After the nodes that \textit{failed\_during\_round} are set to \textit{failed} in line 12, line 13 assumes
- the condition $f<n$. This excludes the cases where all nodes failed.
-
-\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
-\begin{lstlisting}
-  action simulate_failure = {
-    node.failed_during_round(N) := *;
-    assume node.failed_during_round(N) -> ~node.failed(N);
+This section shows how the stopping failure model described by Lynch \cite{refNancy} and explained in \textit{section III ii.} was modeled in Ivy.
+
+To recap, a process may fail before or after performing some instance of Step 1 or Step 2 as well as in the middle of performing Step 1.
+The presented code for actions \textit{step1} and \textit{step2} already correctly considers the behaviour of nodes that failed before the actions.
+E.g.
+\textit{step1} only allows nodes that have not failed to send messages.
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+. . .
+node.incomming_message(O,A) := ~node.failed(A);
+. . .
+\end{lstlisting}
+\end{mdframed}
+
+\subsubsection{Failure before and after rounds}
+
+\noindent Modeling stopping failure before or after any of the steps could therefore be modeled by correctly setting $failed(N)$.
+Adding the following lines before and after \textit{step1} would do the trick:
+
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+node.failing_now(N) := *;
+assume node.failing_now(N) -> ~node.failed(N);
+node.failed_during_round(N) :=
+  (node.failed_now(N) | node.failed_during_round(N))
+node.failed(N) := (node.failed(N) | node.failed_during_round(N));
+assume exists N . ~node.failed(N);
+\end{lstlisting}
+\end{mdframed}
+
+\noindent This snippet is presented only to convey some intuition about the final modeling decision and is not used in the final specification
+
+Line 1 non-deterministically chooses nodes to fail. The \textit{assume} in line 2 excludes all
+cases where a node that already failed was chosen to fail again, as a node can only fail once.
+Line 3 keeps track of the nodes that failed during the current round to enable \textit{clean\_round\_occured} to be set appropriately later.
+Line 5 adds the newly failed nodes to the set of failed nodes.
+Line 6 disallows the possibility that all nodes are chosen to fail and therefore enforces $f < n$.
+
+\subsubsection{Failure during rounds}
+
+However, this specification omits the possibility of node failure during message sending.
+Step 1 is modeled using statements that alter the system state as a whole.
+Expressing non-deterministic failure of individual nodes within these statements is inconvenient.
+The author therefore decided to simulate the behaviour of a node that failed during message sending in a separate piece of code independent of the logic in \textit{step1}.
+Consider the following snippet.
+
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
 node.incomming_message_failing(N,M) := *;
-
 node.incomming_message(N,M) :=
@@
     node.incomming_message(N,M));
-
+\end{lstlisting}
+\end{mdframed}
+
+Line 1 prepares for each node $M$ a non-deterministically chosen set of nodes that it sends a message to.
+This set is represented by the relation \textit{incomming\_message\_failing}.
+This relation can be thought of as the set of nodes $M$ will send a message to if $M$ fails.
+Line 2 updates the \textit{incomming\_message} relation. If node $M$ failed during this round, the set of nodes
+ $M$ sends messages to is changed to the values of \textit{incomming\_message\_failing}, otherwise it remains the same.
+It is important to point out that this approach correctly simulates the fact that nodes are not thought of as sending messages sequentially but in non-deterministic order.
+The specification makes no assumptions about an order of message sending.
+Furthermore, for fixed $m$, \textit{incomming\_message\_failing(N,m)} may still be $true$ for all N.
+The specification therefore includes the possibility of a node failing after it sent all its messages successfully.
+
+\subsubsection{Combination of the two}
+Combining the two principles in a single action gives the following \textit{simulate\_failure} action.
+It is designed to be called once per round between \textit{step1} and \textit{step2}.
+It simulates failure before, during and after \textit{step1}.
+As failure before \textit{step2} is the same as failure after \textit{step1} and failure after \textit{step2} is equivalent to failure before \textit{step1} in the following round, it satisfies all criteria of Lynch's \cite{refNancy} stopping failure model.
+
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+  action simulate_failure = {
+    node.failed_during_round(N) := *;
+    assume node.failed_during_round(N) -> ~node.failed(N);
     node.failed(N) := (node.failed(N) | node.failed_during_round(N));
     assume exists N . ~node.failed(N);
-  }
-\end{lstlisting}
-\end{mdframed}
+
+    node.incomming_message_failing(N,M) := *;
+    node.incomming_message(N,M) :=
+      (node.failed_during_round(M) &
+        node.incomming_message_failing(N,M)) |
+      (~node.failed_during_round(M) &
+        node.incomming_message(N,M));
+
+  }
+\end{lstlisting}
+\end{mdframed}
+
+Lines 2 to 5 resemble the snippet of section \textit{vii.1}.
+In line 2 we choose a random number of nodes to fail.
+The \textit{assume} in line 3 then excludes all cases where a node that already failed was chosen to fail again.
+Line 4 adds the newly failed nodes to the set of failed nodes.
+Line 5 enforces $f < n$.
+Since this action is only called once per round we do not need an extra relation \textit{failing\_now} presented in \textit{vii.1}.
+
+Line 7 to 12 is equivalent to the snippet of \textit{vii.2} and simulates the message sending behaviour of failing nodes.
 

### 2018-08-11 paper/paper.tex
- page 14, section (IV.vi): I would move the explanation about the 	initialization before the sections explaining the steps 1 and 2
- page 15, section (V): remind the reader about the meaning of the correctness 	conditions, and explain them informally before presenting how they are encoded 	in IVy

@@
 
+\subsection{Initialisation}
+
+First, we describe the initial system state in a special \textit{after init} action. When
+proving correctness, IVy requires the initialisation to entail the invariants. All relations and functions
+that are not assigned are set non-deterministically by the environment. E.g. $node.val(N)$ is
+treated as a parameter the environment sets.
+
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+after init {
+  def := *;
+  default(N) := def;
+  w(N, V) := V=node.val(N);
+  incomming_message(A, B) := false;
+  decision(N, V) := false;
+  clean_round_occured := false;
+  decision_round := false;
+  node.failed_before_round(N) := false;
+  node.failed(N) := false;
+  node.has_single_value(N) := false;
+}
+\end{lstlisting}
+\end{mdframed}
+
 \subsection{Rounds}
@@
 
-\subsection{Initialisation}
-
-Finally, we describe the initial system state in a special \textit{after init} action. When
-proving correctness, IVy requires the initialisation to entail the invariants. All relations and functions
-that are not assigned are set non-deterministically by the environment. E.g. $node.val(N)$ is
-treated as a parameter the environment sets.
-
-\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
-\begin{lstlisting}
-after init {
-  def := *;
-  default(N) := def;
-  w(N, V) := V=node.val(N);
-  incomming_message(A, B) := false;
-  decision(N, V) := false;
-  clean_round_occured := false;
-  decision_round := false;
-  node.failed_before_round(N) := false;
-  node.failed(N) := false;
-  node.has_single_value(N) := false;
-}
-\end{lstlisting}
-\end{mdframed}
-
 \section{Verification}
@@
 The simplest invariants to come up with are those, that describe the correctness of the algorithm at hand.
-In our case, these are \textit{agreement}, \textit{validity}, \textit{termination}.
-
-\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
-\begin{lstlisting}
-####### Correctness Conditions #######
-
-# Agreement: No two processes decide on different values.
+In our case, the correctness is described by the \textit{correctness conditions} in \textit{iii}.
+These are \textit{agreement}, \textit{validity} and \textit{termination}.
+
+\subsection{Agreement}
+\textit{Agreement} states that no two correct process decide on different values. It can be described as follows.
+
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
 invariant (~node.failed(A) & ~node.failed(B)) ->
             (node.decision(A,V) = node.decision(B,V))
-
-# every node decides on at most one value
+\end{lstlisting}
+\end{mdframed}
+
+Since the presented specification models deciding via a relation, it is theoretically possible for a node to decide on multiple values.
+To check against this possibility we demand every node may decide on at most one value.
+
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
 invariant node.decision(N,V) -> (~exists O . node.decision(N,O) & O~=V)
-
-# Validity: If all processes start with the same initial value v,
-#           then v is the only possible decision value
+\end{lstlisting}
+\end{mdframed}
+
+\subsection{Validity}
+\textit{Validity} states: If all processes start with the same initial value v, then v is the only possible decision value.
+
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
 invariant (exists N,M . node.val(N) ~= node.val(M)) |
@@
             (~node.failed(N2) -> node.decision(N2,node.val(N2))))
-
-# Termination: All nonfaulty processes eventually decide.
+\end{lstlisting}
+\end{mdframed}
+
+\subsection{Termination}
+\textit{Termination} states, that all non-faulty processes eventually decide.
+
+\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
 invariant system.decision_round <->
@@
 
+\subsection{Invariants for Inductiveness}
 If one were to run $ivy\_check$ on the program without specifying any
@@
 and produces a counter example to induction (CTI). The CTI can be drawn in a GUI.
-In the left hand section of the program we can see, that the system transitioned from state 0
+
+\noindent\includegraphics[width=\textwidth]{gui.png}
+
+\noindent In the left hand section of the program we can see, that the system transitioned from state 0
 to state 1 by executing \textit{round}. In the center IVy displays state 0, meaning the state
@@
 
-\noindent\includegraphics[width=\textwidth]{gui.png}
-
-\noindent State 0 shows, that no node has their initial value, modeled by $val$,
+State 0 shows, that no node has their initial value, which is modeled by the relation $val$,
 in its $w$. The user must now realise, that this is an unreachable state, and exclude it by coming
@@
 
-The method for finding the remaining invariants is similar: Execute $ivy\_check$ $\rightarrow$ examine CTI $\rightarrow$ design invariant $\rightarrow$ execute $ivy\_check$.
+The method for finding the remaining invariants is identical: Execute $ivy\_check$ $\rightarrow$ examine CTI $\rightarrow$ design invariant $\rightarrow$ execute $ivy\_check$.
+The following snippets presents the necessary invariants for induction annotated with comments.
 \begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
@@
 invariant node.default(N) = node.def
+
 # if a value is in a process' W, it must be initial value of
 # some other process (no invented values)
+
 invariant node.w(N, V) -> (exists M . node.val(M)=V)
+
 # lemma 1:

### 2018-08-11 paper/paper.tex
nobreak mdframed
proof -> prove

@@
 It takes as input a specification of an algorithm operating on system-state  and a set of invariants.
-It proofs the inductiveness of the invariants in respect to the algorithm, or displays a counterexample to induction.
+It proves the inductiveness of the invariants in respect to the algorithm, or displays a counterexample to induction.
 If the set of invariants entails the desired properties of the algorithm, and Ivy proved induction, then it was proven that the algorithm is \textit{correct}.
@@
 For example to verify servers that specify connection and disconnection algorithms.
-This can be thought of as proofing inductiveness of two independent algorithms with respect to the same invariants and system-state.
+This can be thought of as proving inductiveness of two independent algorithms with respect to the same invariants and system-state.
 
@@
 In IVy one mainly uses \textit{uninterpreted} types. \textit{Uninterpreted} means that the type is an arbitrary set of values with at least one element.
-The cardinality of this set is \textit{parameterized}, meaning IVy proofs correctness for an arbitrary amount of values of an \textit{uninterpreted} type.
+The cardinality of this set is \textit{parameterized}, meaning IVy proves correctness for an arbitrary amount of values of an \textit{uninterpreted} type.
 The number of values of a type that are generated to construct a counter example to induction (CTI) is entirely up to IVy. No instances of these types are manually instantiated.
@@
 
-\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
 \begin{lstlisting}
@@
 To assign values to relations, functions or variables $:=$ is used.
-\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
 \begin{lstlisting}
@@
 \subsubsection{Invariants}
-To specify the properties we want to proof about our algorithm we encode them in special assertions called \textit{invariants}:
-
-\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+To specify the properties we want to prove about our algorithm we encode them in special assertions called \textit{invariants}:
+
+\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
 \begin{lstlisting}
@@
 
-We refer to these condition as ``correctness conditions'', as we say the algorithm is ``correct'' if we can proof it satisfies these conditions.
+We refer to these condition as ``correctness conditions'', as we say the algorithm is ``correct'' if we can prove it satisfies these conditions.
 \textit{Floodset} solves the described problem by defining $msgs$ and $trans$ functions for nodes in the network such that the conditions are satisfied.
@@
 \textbf{Formal definition of \textit{Floodset}}:
-\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
 \noindent $states_i$:
@@
 
-\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
 \noindent $msgs_i$:
@@
 
-\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
 \noindent $trans_i$:
@@
 It is trivial to verify for a small number of $n,\ f$, but the question is ``does \textit{Floodset} work for any assignment of $n,\ f,\ n < f$''.
-To answer this question the algorithm was encoded in the IVy language and then inductive invariants were constructed to proof correctness.
+To answer this question the algorithm was encoded in the IVy language and then inductive invariants were constructed to prove correctness.
 
@@
 IVy therefore allows specifying invariants together with a loop.
-\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
 \begin{lstlisting}
@@
 \noindent A round in the specification is defined as follows:
-\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
 \begin{lstlisting}
@@
 \textit{step1} only allows nodes that have not failed to send messages.
-\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
 \begin{lstlisting}
@@
 The invariant excluding the unreachable state says, that the initial value of a node is always in its $w$:
-\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
 \begin{lstlisting}
@@
 The following snippets presents the necessary invariants for induction annotated with comments.
-\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
 \begin{lstlisting}
@@
 
-How viable is IVy as a tool for algorithm designers to proof correctness
+How viable is IVy as a tool for algorithm designers to prove correctness
 of their work?  The goal, when turning to a tool like IVy, is certainty of
@@
 
-as
-
-\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\noindent as
+
+\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
 \begin{lstlisting}
@@
 insert
-\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
 \begin{lstlisting}
@@
 in favor of the logic oriented language. It gave an explanation of induction and inductive invariants,
-explaining what they are, how they are used by IVy to proof correctness and how they can be found.
+explaining what they are, how they are used by IVy to prove correctness and how they can be found.
 It presented some of the systems capabilities to verify distributed algorithms by example.

### 2018-08-12 paper/paper.tex
- page 16, section (V): give an intuitive explanation about the invariants 	imposed by the framework. You could also pick one and explain it in plain 	English

@@
 
-Since it is also required that the \textit{framework} behaves correctly it makes sense to include some
-correctness conditions specifically aimed at the system.
+\subsection{Checking framework correctness}
+
+Since any check on \textit{Floodset}'s correctness assumes that the \textit{framework} has been modeled correctly, it makes sense to include correctness conditions specifically aimed at the \textit{framework} in the form of invariants.
+These invariants are not necessary for induction.
+Removing them will not cause any of Ivy's other checks to fail.
+Their sole purpose is to give the user confidence that the \textit{framework} is specified correctly.
+
+For example, we would like to check that $f < n$ at all times.
+This has nothing to do with \textit{Floodset}'s correctness and does not have to be specified for inductiveness.
+However, to the user it is useful to make sure that this condition is never broken.
+Line 4 checks $f < n$ by using the previously presented alternate encoding using an existential quantifier.
+
+This snippet also shows some other simple checks of this kind. They are annotated with comments to facilitate understanding of their purpose.
 
@@
 \end{mdframed}
+
+These invariants do not constitute a full-fledged proof of the framework being correctly modeled.
+For example, we have not proven that the \textit{framework} ever fails a node.
+This limitation will be discussed in detail in the next section.
 

### 2018-08-12 paper/paper.tex
- I would also mention that the properties checked by IVy are safety 	properties. Checking liveness is also possible, but one would need to 	construct a reduction of liveness to safety [2] which is out of the scope of 	this work

@@
 \end{mdframed}
+This invariant assumes that the system will eventually reach the decision round.
+However, truly verifying \textit{termination} would require checking that the system must reach the decision round, which means verifying  \textit{liveness}.
+Ivy theoretically allows \textit{liveness} checks by constructing a reduction of \textit{liveness} to \textit{safety} \cite{live} but this is out of the scope of this work.
 
@@
     \bibitem{ivy} O. Padon, K. McMillan, A. Panda, M. Safiv, S. Shoham, ``Ivy: Safety Verification by Interactive Generalization'', Available \textit{https://www.cs.tau.ac.il/~odedp/ivy.pdf}
+    \bibitem{live} O.
+Padon, J.
+Hönicke, G.
+Losa, A.
+Podelski, M.
+Sagiv, S.
+Shoham, ``Reducing Liveness to Safety in First-Order Logic'', Available \textit{https://www.cs.tau.ac.il/~odedp/reducing-liveness-to-safety-in-first-order-logic/}
     \bibitem{refLanguageDoc} ``The IVy language'',  \textit{http://microsoft.github.io/ivy/language.html} [Accessed: 5.06.18], 2017
@@
     \bibitem{github} C. Claus, ``BA'', \textit{https://github.com/calvinclaus/ba}, 2018
+
 \end{thebibliography}

### 2018-08-12 paper/paper.tex
- page 17, section (VI): I think the message of this section is that one needs 	to have good understanding of the algorithm modeled, and the conditions under 	which it is supposed to work. Maybe you could add a bit more structure to it 	to make it more clear. Although the implementation is not verified, the goal 	is to verify the actual design of the algorithm before it is ever implemented, 	and this is where IVy comes useful. Also, I do not completely get what do you 	mean by "if would be beneficial to have a built in mechanism to verify the 	possibility of something to occur".    If you find

@@
 \section{Evaluation}
-
-How viable is IVy as a tool for algorithm designers to prove correctness
+How viable is IVy as a tool for algorithm designers proving correctness
 of their work? The goal, when turning to a tool like IVy, is certainty of
-correctness. However, absolute certainty is never really achieved.
-One can not trivially determine, that the specification in the IVy language is logically equivalent to the actual algorithm.
-This is compounded by the fact, that the IVy language is structurally unlike classical procedural languages. Making it challenging to achieve
-a one-to-one translation from a classical distributed algorithm in a language like C. An example of this is
-the relatively complicated quantifier-logic in assignments that update the global system state.
-
-Another threat is posed by $assume$ statements. The author of this paper originally modeled the line
+correctness. However, due to a variety of reasons, absolute certainty is never achieved.
+This section evaluates Ivy's usefulness pointing out strengths and limitations that Ivy users should keep in mind
+when working with the system.
+
+\subsection{Equivalence of specification and algorithm}
+One can not trivially determine, that the specification in the IVy language is logically equivalent to the actual algorithm the user wishes to verify.
+Whether the algorithm to be verified exists in the user's mind, in the form of pseudo-code or fully implemented in a programming language, no guarantee can be given in regards to the correctness of the translation of the algorithm into the Ivy language.
+Some examples illustrating this difficulty will be given in the following paragraphs.
+
+\subsubsection{Structural difference of Ivy and procedural languages}
+The IVy language is structurally unlike classical procedural languages.
+This makes it inherently challenging to achieve a one-to-one translation from a classical distributed algorithm, that was specified in a procedural way, like \textit{Floodset}.
+An example of this structural difference is the relatively complicated quantifier-logic in assignments that update the global network state.
+Take as an example the first line of \textit{step2}:
+
+\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+node.w(N,V) := (node.w(N,V) |
+               (~node.failed(N) &
+                exists J . (node.incomming_message(N,J) &
+                             node.message(J,V))));
+\end{lstlisting}
+\end{mdframed}
+
+\noindent which models
+
+\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{gather*}
+  \mathrm{let}\ X_j\ \mathrm{be\ the\ message\ from\ j,\ for\ each\ j\ from\ which\ a\ message\ arrives}\\
+  W\ :=\ W \cup \bigcup_{j} X_j\\
+\end{gather*}
+\end{mdframed}
+
+This comparison clearly shows the complex structural differences.
+While the pseudo-code is concerned only with state alteration of a single node, the Ivy language code simultaneously alters the state of all nodes.
+It uses an existential quantifier nested within a term.
+Its logic touches on multiple layers of abstraction as it mixes system and node concerns.
+It considers node failure, message reception and node state alteration.
+
+As these two lines are so different, equivalence is certainly not obvious to a user.
+Proving equivalence of pseudo-code and Ivy language specification would certainly be a non-trivial proof exercise that certainly trumps manually proving \textit{Floodset}'s correctness in difficulty.
+
+\subsubsection{Lack of modularity}
+Furthermore, it has already been shown in section \textbf{IV.
+i.1} that the lack of easy iteration over types makes it difficult to achieve modularity.
+The lack of which introduces additional uncertainty about correct modelling, especially in distributed systems.
+
+\subsection{Accidental invalidation of legal state configurations with \textit{assume}}
+$assume$ statements may silently invalidate state configurations. They therefore pose a threat as they
+may allow false-positive Ivy checks.
+To give a practical example, the author of this paper originally modeled the line
 
@@
 \end{mdframed}
- and did not discover that the $assume$ is bogus and excludes important possibilities from being verified.
+The second snippet misses an existential quantifier on line 4, and therefore specifies behaviour that is unlike that of \textit{Floodset}.
+The author did not discover that this $assume$ is bogus and excludes valid system states from being verified.
 The error only became apparent when it was discovered that correctness could be proven in the absence of an invariant that should have been absolutely necessary for induction.
 
-Furthermore, it has already been shown, that the lack of easy iteration over types makes it difficult to achieve modularity.
-The lack of which introduces additional uncertainty about correct modelling, especially in distributed systems.
-
-However, as has been pointed out, it is not just the algorithm that must be correctly specified, but the underlying \textit{framework} is equally important to get right.
+\subsection{Miss-specification of system context}
+As has been pointed out, it is not just the algorithm that must be correctly specified.
+The underlying \textit{framework} is equally important to get right.
 In the presented specification it is possible to remove the \textit{simulate\_failure} action and still receive a positive check from the IVy system.
-This is unsurprising, as we do not have any mechanism in place that checks that the system correctly models the possibility for node failure.
-E.g.
-to check that the simulate failure action allows a node to fail and send a message only to a subset of other nodes one can purposefully insert an assertion that only fails when the desired behaviour is possible, e.g.
-insert
+This is unsurprising, as we do not have any mechanism in place verifying that the system allows nodes to fail.
+More subtly, a user may incorrectly simulate partial message delivery of failing nodes, however all checks would still pass.
+
+In general, the user is left with the question ``does my \textit{framework}-specification correctly allow for X to happen''.
+Users may therefore want to verify, that their specification allows a certain state to occur.
+In regard to Lynch's \textit{framework} the user may wish to check that a node can fail during message delivery, resulting in partial message delivery.
+
+The author of this paper used the following approach to find out if he encoded node failure correctly:
+\begin{enumerate}
+  \item Construct an assertion describing the system state that should be allowed to occur
+  \item Negate the assertion
+  \item Add the assertion to an appropriate location in the specification
+  \item Check that assertion fails
+  \item Remove assertion
+\end{enumerate}
+
+\noindent For example, one may insert
 \begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
@@
 after the \textit{simulate\_failure} action.
-If IVy does not find a counterexample to this assert, the specification is erroneous.
-In light of this, it would be beneficial to have a built in mechanism to verify the possibility of something to occur.
+The assertion checks that a node cannot fail and not deliver a message to J while successfully delivering a message to K.
+If IVy does not find a counterexample to this assertion, it means that the specification does not allow partial message delivery.
+The specification would therefore be erroneous.
+One has to finally remove the failing assertion to allow Ivy to proof induction again.
+
+In light of this, it would be beneficial to have a built in mechanism to verify the possibility of states to occur.
 In this case, it would allow one to verify, that the modeled system allows for the possibility of a node to fail such that it only sends messages to a subset of processes, as part of the final proof.
 
-It has become apparent, that successful verification with IVy is only possible with deep knowledge about how the algorithm functions. It is not a ``translate and go'' solution.
+Of course this approach does not constitute a proof of correct \textit{framework}-specification. It solely acts as a form of test for the user.
+
+\subsection{Necessity of deep understanding}
+It has become apparent, that successful verification with IVy is not possible without deep knowledge about how the algorithm functions. Ivy is not a ``translate and go'' solution.
 Evidence of this include the complex invariant that requires knowledge about \textit{Lemma 1} or the alternative encoding for $f<n$ and the decision condition $if\ rounds = f+1$ that had to be discovered.
 
-Of course, a variety of bugs and exploits do not stem from a logically unsound algorithmic foundation but implementation errors.
+\subsection{Algorithm failure in theory and practice}
+As a user it is important to understand that Ivy's goal is to verify algorithmic correctness on a high level.
+It does not concern itself with implementation details.
+However, a variety of bugs and exploits do not stem from logically unsound algorithmic foundations but implementation errors.
 A wonderful example is the Heartbleed exploit in OpenSSL.
-It would have been no help to verify the simple Heartbleed protocol in IVy before implementing it in C, since the error was in an implementation detail allowing a buffer underflow.
-The point is, the implementation of the algorithm must be tested through unit testing or bounded model checkers, despite successful verification in IVy.
-
+It would have been no help to verify the Heartbeat protocol in IVy before implementing it in C, since the error was in an implementation detail allowing a buffer underflow.
+The Ivy user must realise, that the implementation of the algorithm must be tested through unit testing and/or bounded model checkers, despite successful verification in IVy.
+
+\subsection{Time of specification}
+Unfortunately the author missed the opportunity to scientifically record the time commitment of verifying \textit{Floodset} in Ivy.
+He estimates the time from starting to learn Ivy to successfully proving \textit{Floodset} at around 150 hours.
+In the case of \textit{Floodset} manual proving would have certainly been significantly faster.
+Drawing a general conclusion about practicability of Ivy is not possible from this data.
+
+For future work, it may be interesting to study whether specification in Ivy gets easier or harder than manual proving as algorithmic complexity increases.
+
+\subsection{User friendliness}
 Despite these limitations, IVy trumps similar tools in user friendliness by upping interactivity. \cite{ivy}
@@
 
-Therefore, if a user is capable of navigating the limitations with care, IVy is absolutely adequate for proving the logical soundness of
-an algorithm.
+Therefore, if a user is capable of navigating the limitations with care, IVy is absolutely adequate for proving the logical soundness of an algorithm.
 

### 2018-08-12 paper/paper.tex
Improved explanation of method to verify a system state can ocur

@@
 
-The author of this paper used the following approach to find out if he encoded node failure correctly:
+One may use the following approach to verify that a particular system state is reachable:
 \begin{enumerate}
-  \item Construct an assertion describing the system state that should be allowed to occur
-  \item Negate the assertion
-  \item Add the assertion to an appropriate location in the specification
-  \item Check that assertion fails
-  \item Remove assertion
+  \item Construct a formula, $\gamma$, describing the system state in question.
+  \item Negate the formula, $\neg\gamma$.
+  \item Add \textit{assert} $\neg\gamma$ to an appropriate location in the specification.
+  \item Check that the assertion fails.
+  \item Remove the assertion.
 \end{enumerate}
@@
 \begin{lstlisting}
-ensure ~(node.failed(N) &
+assert ~(exists N . node.failed(N) &
         (exists J . ~incoming_message(J,N) &
@@
 after the \textit{simulate\_failure} action.
-The assertion checks that a node cannot fail and not deliver a message to J while successfully delivering a message to K.
-If IVy does not find a counterexample to this assertion, it means that the specification does not allow partial message delivery.
+The negated part, $\gamma$, specifies that there exists a node $N$, that failed and failed to deliver a message to $J$ while successfully delivering a message to $K$, with $N,J,K$ distinct.
+By negating $\gamma$, we force Ivy to construct a counterexample such that $\gamma$ is satisfied.
+If IVy does not find a counterexample to this assertion, it means that $\gamma$ cannot occur.
+In this particular case, that would mean the specification does not allow partial message delivery.
 The specification would therefore be erroneous.
-One has to finally remove the failing assertion to allow Ivy to proof induction again.
-
-In light of this, it would be beneficial to have a built in mechanism to verify the possibility of states to occur.
+One has to finally remove the failing assertion to allow Ivy to prove induction again.
+
+In light of this, it would be beneficial to have a built in mechanism to verify the possibility of certain states to occur.
 In this case, it would allow one to verify, that the modeled system allows for the possibility of a node to fail such that it only sends messages to a subset of processes, as part of the final proof.
 
-Of course this approach does not constitute a proof of correct \textit{framework}-specification. It solely acts as a form of test for the user.
+Of course this approach would not constitute a proof of correct \textit{framework}-specification. It solely acts as a form of test.
 

### 2018-08-12 paper/paper.tex
- page 1, introductory paragraph: "not to be satisfied" can be replaced by

@@
 The environment conceptually represents the final users of the algorithm.
-An algorithm or protocol is \textit{correct} or \textit{safe} if it cannot be called by the environment in a state configuration that causes some pre-defined properties (assertions, invariants) not to be satisfied.
+An algorithm or protocol is \textit{correct} or \textit{safe} if it cannot be called by the environment in a state configuration that causes some pre-defined properties (assertions, invariants) to be violated.
 IVy supports verification of systems that have multiple methods (\textit{actions}) exported to the environment.

### 2018-08-12 paper/paper.tex
- page 1, introduction: it's -> it is, it's -> its
- page 2, section (II.i.3): "As a user ..." -> "To a user ..."

@@
 \end{enumerate}
-As a user without any experience in automatic verification, \textit{inductive} invariants are a difficult concept.
+To a user without any experience in automatic verification, \textit{inductive} invariants are a difficult concept.
 IVy does not consider the reachability of a system state.  As long as the state is in compliance with the invariants it is assumed to be reachable.
@@
 E.g.
-for \textit{Step 1} the system would iterate over all nodes and let each node take care of sending it's own message.
+for \textit{Step 1} the system would iterate over all nodes and let each node take care of sending its own message.
 

### 2018-08-12 paper/paper.tex
- page 3, section (II.ii): "... the language through which *the behavior of* 	the algorithm is defined"
- page 3, section (II.ii.4): "*For example, * this line 	non-deterministically..."

@@
 \subsection{The IVy language}
-The IVy language is a core part of the IVy system. It is the language through which the algorithm to be verified is defined.
+The IVy language is a core part of the IVy system. It is the language through which the behaviour of the algorithm to be verified is defined.
 This section presents a selection of core language features that should help the reader understand the paradigm of modelling in IVy, as well as lay the necessary foundation to understand
@@
 To non-deterministically assign to a variable, relation or function ``*'' is used.
-This line non-deterministically chooses a number of nodes to have failed.
+For example, this line non-deterministically chooses a number of nodes to have failed.
 

### 2018-08-12 paper/paper.tex
clearity
- page 6, section (III): "Lynch's work" -> "Lynch's book" 	- page 9, section (III.iii): Thorem -> Theorem

@@
 
-The algorithm that was implemented in IVy as part of this work is called \textit{Floodset} and was taken out of Lynch's work ``Distributed Algorithms'' \cite{refNancy}.
+The algorithm that was implemented in IVy as part of this work is called \textit{Floodset} and was taken out of Lynch's book ``Distributed Algorithms'' \cite{refNancy}.
 \textit{Floodset} is a consensus algorithm for synchronous networks with failing processes.
@@
 \subsection{Stopping failures}
-The system presented in Lynch's work \cite{refNancy} is capable of modelling failure in a variety of ways.
+The system presented in Lynch's book \cite{refNancy} is capable of modelling failure in a variety of ways.
 This work makes use of stopping failure only. This kind of failure occurs when a process stops in the middle of its execution. With regards to the presented \textit{framework} this
@@
 
-\noindent What follows is a proof of the correctness of the \textit{Floodset} algorithm to illustrate the principles behind it.
+\noindent What follows is a proof of the correctness of the \textit{Floodset} algorithm to illustrate the principles behind the algorithm.
 It has already been proven correct in \cite{refNancy}.
@@
 \noindent The Lemmas allow proving the theorem:\\
-\noindent \textbf{Thorem}:\\
+\noindent \textbf{Theorem}:\\
 \textit{Floodset} satisfies the correctness conditions.\\

### 2018-08-12 paper/paper.tex
- page 10, section (IV.i.2): couldn't -> could not

@@
 When programming in a procedural language one would use a loop for this task but, as has been explained in the previous section, loops are somewhat problematic in IVy.
-The author couldn't find a way to specify an inductive loop invariant for a loop counting the number of failed nodes without a cardinality operator at hand.
+The author could not find a way to specify an inductive loop invariant for a loop counting the number of failed nodes without a cardinality operator at hand.
 This means that alternative encodings for the assumption $f<n$ and the \textit{decision condition} ($if\ rounds = f+1$) must be found.

### 2018-08-12 paper/paper.tex
- page 13, section (IV.iii): don't -> do not, incomming_message -> 	incoming_message, hasn't -> has not

@@
 Therefore the actions will choose a random non-failed node to send a message to x.
-If there is no node that hasn't failed, this action cannot terminate.
+If there is no node that has not failed, this action cannot terminate.
 
@@
   node.message(n,V) := node.w(n,V);
-  node.incomming_message(O,n) := true;
+  node.incoming_message(O,n) := true;
 }
@@
 line 3: message := *;
-line 3: incomming_message := *;
+line 3: incoming_message := *;
 . . .
@@
   invariant node.message(N,V) = node.w(N,V)
-  invariant N <= iter.val -> incomming_message(O,N)
+  invariant N <= iter.val -> incoming_message(O,N)
   {
@@
 assert message(N,V) = w(N,V)
-assert N:t <= t.iter.val(loc:iter) -> incomming_message(O,N)
-incomming_message := *
+assert N:t <= t.iter.val(loc:iter) -> incoming_message(O,N)
+incoming_message := *
 message := *
 assume message(N,V) = w(N,V)
-assume N:t <= t.iter.val(loc:iter) -> incomming_message(O,N)
+assume N:t <= t.iter.val(loc:iter) -> incoming_message(O,N)
 \end{lstlisting}
@@
 
-  # incomming_message(n,o) denotes whether node n received a message
+  # incoming_message(n,o) denotes whether node n received a message
   # from node o
-  relation incomming_message(SELF:node, OTHER:node)
+  relation incoming_message(SELF:node, OTHER:node)
 
@@
   # helper to simulate non-determinstic message sending
-  relation incomming_message_failing(SELF:node, OTHER:node)
+  relation incoming_message_failing(SELF:node, OTHER:node)
 
@@
   w(N, V) := V=node.val(N);
-  incomming_message(A, B) := false;
+  incoming_message(A, B) := false;
   decision(N, V) := false;
@@
 $message(1,0) = true; message(1,1) = true;$ means node $1$ includes in its message the values $0$ and $1$.
-Since a node sends the same message to all nodes, for any given round, we don't need to include the
+Since a node sends the same message to all nodes, for any given round, we do not need to include the
 recipient in the relation.
 
-The actual delivery of a message is modeled via the $incomming\_message(RECIPIENT:node,SENDER:node)$ relation.
+The actual delivery of a message is modeled via the $incoming\_message(RECIPIENT:node,SENDER:node)$ relation.
 Where $incoming\_message(0,1) = true; incoming\_message(0,2) = true;$ means that node $0$ receives the messages from node $1$ and node $2$.
@@
 Line 2 assigns for all nodes $A$ and values $V$ to $message(A,V)$ the boolean value of $node.w(A,V) \land \neg failed(A)$.
-In other words, every node that hasn't failed prepares a message with all values in its $w$.
+In other words, every node that has not failed prepares a message with all values in its $w$.
 Line 3 states that all nodes that haven't failed send messages to all other nodes.
@@
   node.message(A,V) := node.w(A,V) & ~node.failed(A);
-  node.incomming_message(O,A) := ~node.failed(A);
+  node.incoming_message(O,A) := ~node.failed(A);
 }
@@
 they decide on the default value.
-Finally, for good measure, $incomming\_message$ and $message$ are reset, but this is not necessary for the proof.
+Finally, for good measure, $incoming\_message$ and $message$ are reset, but this is not necessary for the proof.
 
@@
                  (~node.failed(N) &
-                  exists J . (node.incomming_message(N,J) &
+                  exists J . (node.incoming_message(N,J) &
                                node.message(J,V))));
@@
 
-  node.incomming_message(A,B) := false;
+  node.incoming_message(A,B) := false;
   node.message(A,V) := false;
@@
 . . .
-node.incomming_message(O,A) := ~node.failed(A);
+node.incoming_message(O,A) := ~node.failed(A);
 . . .
@@
 \begin{lstlisting}
-node.incomming_message_failing(N,M) := *;
-node.incomming_message(N,M) :=
+node.incoming_message_failing(N,M) := *;
+node.incoming_message(N,M) :=
   (node.failed_during_round(M) &
-    node.incomming_message_failing(N,M)) |
+    node.incoming_message_failing(N,M)) |
   (~node.failed_during_round(M) &
-    node.incomming_message(N,M));
+    node.incoming_message(N,M));
 \end{lstlisting}
@@
 Line 1 prepares for each node $M$ a non-deterministically chosen set of nodes that it sends a message to.
-This set is represented by the relation \textit{incomming\_message\_failing}.
+This set is represented by the relation \textit{incoming\_message\_failing}.
 This relation can be thought of as the set of nodes $M$ will send a message to if $M$ fails.
-Line 2 updates the \textit{incomming\_message} relation. If node $M$ failed during this round, the set of nodes
- $M$ sends messages to is changed to the values of \textit{incomming\_message\_failing}, otherwise it remains the same.
+Line 2 updates the \textit{incoming\_message} relation. If node $M$ failed during this round, the set of nodes
+ $M$ sends messages to is changed to the values of \textit{incoming\_message\_failing}, otherwise it remains the same.
 It is important to point out that this approach correctly simulates the fact that nodes are not thought of as sending messages sequentially but in non-deterministic order.
 The specification makes no assumptions about an order of message sending.
-Furthermore, for fixed $m$, \textit{incomming\_message\_failing(N,m)} may still be $true$ for all N.
+Furthermore, for fixed $m$, \textit{incoming\_message\_failing(N,m)} may still be $true$ for all N.
 The specification therefore includes the possibility of a node failing after it sent all its messages successfully.
@@
 
-    node.incomming_message_failing(N,M) := *;
-    node.incomming_message(N,M) :=
+    node.incoming_message_failing(N,M) := *;
+    node.incoming_message(N,M) :=
       (node.failed_during_round(M) &
-        node.incomming_message_failing(N,M)) |
+        node.incoming_message_failing(N,M)) |
       (~node.failed_during_round(M) &
-        node.incomming_message(N,M));
+        node.incoming_message(N,M));
 
@@
 # 'after' a node has failed it will not send anything
-invariant node.failed_before_round(N) -> ~node.incomming_message(M, N)
+invariant node.failed_before_round(N) -> ~node.incoming_message(M, N)
 
@@
                (~node.failed(N) &
-                exists J . (node.incomming_message(N,J) &
+                exists J . (node.incoming_message(N,J) &
                              node.message(J,V))));
@@
 assume node.w(A,V) <-> (node.tmp_w(A,V) |
-                        ((node.incomming_message(A,J) &
+                        ((node.incoming_message(A,J) &
                           node.message(J,V))));
@@
         (exists J . ~incoming_message(J,N) &
-          exists K . incomming_message(K,N) & J~=N & K~=J & N~=K))
+          exists K . incoming_message(K,N) & J~=N & K~=J & N~=K))
 \end{lstlisting}

### 2018-08-12 paper/paper.tex
Minor fixes

@@
 IVy is a system designed to verify correctness of infinite-sate systems using the aforementioned deductive approach.
-It takes as input a specification of an algorithm operating on system-state  and a set of invariants.
+In a nutshell, it takes as input the specification of an algorithm, which may mutate state, and a set of invariants.
 It proves the inductiveness of the invariants in respect to the algorithm, or displays a counterexample to induction.
@@
 
-A detailed discussion about decidability lies outside of the scope of this work and is well documented in \cite{decide}.
+A detailed discussion about decidability lies outside of the scope of this work and is well documented in \cite{decid}.
 
@@
 
-The assumption $f<n$, ``number of failed nodes is bounded to an amount smaller than the number of total nodes'', can alternatively be expressed via $\exists n\ \neg failed(N)$.
+The assumption $f<n$, ``number of failed nodes is bounded to an amount smaller than the total number of nodes'', can alternatively be expressed via $\exists n\ \neg failed(n)$.
 This is because the existence of a functioning node implies that the number of failed nodes $f$ is smaller than $n$.
@@
 The contents of a node's message is modeled via the relation $message(N:node,V:value)$. Where
-$message(1,0) = true; message(1,1) = true;$ means node $1$ includes in its message the values $0$ and $1$.
+$message(1,0) = true;\ message(1,1) = true;$ means node $1$ includes in its message the values $0$ and $1$.
 Since a node sends the same message to all nodes, for any given round, we do not need to include the
@@
 
-The actual delivery of a message is modeled via the $incoming\_message(RECIPIENT:node,SENDER:node)$ relation.
-Where $incoming\_message(0,1) = true; incoming\_message(0,2) = true;$ means that node $0$ receives the messages from node $1$ and node $2$.
+The actual delivery of a message is modeled via the \\
+$incoming\_message(RECIPIENT:node,SENDER:node)$ relation. Where $incoming\_message(0,1) = true;\ incoming\_message(0,2) = true;$ means that node $0$ receives the messages from node $1$ and node $2$.
 
@@
 
-Line 2 in the \textit{step2} action adds to $w$ of node $N$ all values $V$ for which a node $J$ exists that has sent a message to $N$ that contains $J$.
-The conditional statement in Line 4 corresponds to the \textit{decision condition} ($if\ rounds = f+1$) in the original algorithm. Nodes with a single value in $w$ decide on that value, otherwise
-they decide on the default value.
+Line 2 in the \textit{step2} action adds to $w$ of node $N$ all values $V$ for which a node $J$ exists that has sent a message to $N$ that contains $V$.
+The conditional statement in Line 7 corresponds to the \textit{decision condition} ($if\ rounds = f+1$) in the original algorithm.
+Line 9 sets the helper relation $has\_single\_value(N:node)$ that is set to true for $N$ iff it has exactly one value in its $w$.
+Line 13 specifies that nodes with a single value in $w$ decide on that value, unless they have failed.
+Line 16 specifies that nodes with multiple values in their $w$ decide on their default value, unless they failed.
+The disjunction on line 17 with $node .
+decision (N, node .
+default (N))$ is necessary so as not to overwrite an already made decision of a node with $w = \{v_0\}$, $v_0$ being the default value.
 Finally, for good measure, $incoming\_message$ and $message$ are reset, but this is not necessary for the proof.
@@
 This set is represented by the relation \textit{incoming\_message\_failing}.
-This relation can be thought of as the set of nodes $M$ will send a message to if $M$ fails.
-Line 2 updates the \textit{incoming\_message} relation. If node $M$ failed during this round, the set of nodes
+This relation can be thought of as the set of nodes which $M$ will send a message to if $M$ fails.
+Line 2 updates the \textit{incoming\_message} relation. If node $M$ failed during this round, the set of nodes which
  $M$ sends messages to is changed to the values of \textit{incoming\_message\_failing}, otherwise it remains the same.
@@
 The simplest invariants to come up with are those, that describe the correctness of the algorithm at hand.
-In our case, the correctness is described by the \textit{correctness conditions} in \textit{iii}.
+In our case, the correctness is described by the \textit{correctness conditions} in \textit{III.iii}.
 These are \textit{agreement}, \textit{validity} and \textit{termination}.
@@
 # some other process (no invented values)
-
 invariant node.w(N, V) -> (exists M . node.val(M)=V)
 
-# lemma 1:
+# lemma 1, explained in III
 invariant system.clean_round_occured ->
@@
 
-As these two lines are so different, equivalence is certainly not obvious to a user.
-Proving equivalence of pseudo-code and Ivy language specification would certainly be a non-trivial proof exercise that certainly trumps manually proving \textit{Floodset}'s correctness in difficulty.
+As these two lines are so different, equivalence may not be obvious to a user.
+Proving equivalence of pseudo-code and Ivy language specification would be a non-trivial proof exercise that certainly trumps manually proving \textit{Floodset}'s correctness in difficulty.
 
 \subsubsection{Lack of modularity}
-Furthermore, it has already been shown in section \textbf{IV.
+Furthermore, it has already been shown in section \textit{IV.
 i.1} that the lack of easy iteration over types makes it difficult to achieve modularity.
@@
 It has become apparent, that successful verification with IVy is not possible without deep knowledge about how the algorithm functions. Ivy is not a ``translate and go'' solution.
-Evidence of this include the complex invariant that requires knowledge about \textit{Lemma 1} or the alternative encoding for $f<n$ and the decision condition $if\ rounds = f+1$ that had to be discovered.
+Evidence of this include the complex invariant that requires knowledge about \textit{lemma 1} or the alternative encoding for $f<n$ and the decision condition $if\ rounds = f+1$.
 
@@
 This paper has presented the Ivy system and important Ivy language features. It classified IVy
-as part of a three step process of algorithm verification.
+as part of a two step process of algorithm verification.
 It explained that its usage requires a readiness to let go of procedural programming paradigms
@@
 
-It pointed out some limitations. One being that it is hard to know an algorithm was correctly specified in IVy. This is due to
-the lack of easy type iteration, which leads to complicated global assignments and a non-modular structure that is far off
-the original algorithm's.
-The paper suggested a mechanism to verify the possibility of a state to occur, to alleviate this uncertainty.
+It pointed out some limitations.
+One being that it is hard to know an algorithm was correctly translated into the IVy language, as a one-to-one translation from a procedural algorithm is unlikely.
+This is due to the lack of easy type iteration, which leads to complicated global assignments and a non-modular structure that is far off the original algorithm's.
+Furthermore, the system context may easily be miss-specified. The paper suggested a mechanism to verify the possibility of a state to occur, to alleviate this problem.
 Another limitation that was found, is the difference between a logically sound algorithm and a correctly implemented one, and how IVy can only verify the former.

### 2018-08-22 paper/paper.tex
implented comments2

@@
 
-  Ivy is tailored to simplify deductive verification of parameterized distributed algorithms.
+  IVy is tailored to simplify deductive verification of parameterized distributed algorithms.
 There are a variety of systems currently implementing the deductive verification process.
@@
 It proves the inductiveness of the invariants in respect to the algorithm, or displays a counterexample to induction.
-If the set of invariants entails the desired properties of the algorithm, and Ivy proved induction, then it was proven that the algorithm is \textit{correct}.
-
-In Ivy, the entity ``calling'' the algorithm in different state configurations is called the ``environment''.
+If the set of invariants entails the desired properties of the algorithm, and IVy proved induction, then it was proven that the algorithm is \textit{correct}.
+
+In IVy, the entity ``calling'' the algorithm in different state configurations is called the ``environment''.
 The environment conceptually represents the final users of the algorithm.
@@
 \subsubsection{Step 1 - encoding in IVy language}
-Step 1 requires the user to understand the Ivy language.
+Step 1 requires the user to understand the IVy language.
 This language is a specification language tailored to describe state changes in a system.
-Ivy language code is inherently different from procedural languages, as it serves a different purpose and its logic is never executed.
-The Ivy documentation itself calls the language features "unusual" \cite{refLanguageDoc} so a user may need considerable effort to understand how to best use its tools.
-
-Using their understanding of the Ivy language users have to encode their algorithm in it.
+IVy language code is inherently different from procedural languages, as it serves a different purpose and its logic is never executed.
+The IVy documentation itself calls the language features "unusual" \cite{refLanguageDoc} so a user may need considerable effort to understand how to best use its tools.
+
+Using their understanding of the IVy language users have to encode their algorithm in it.
 However, distributed algorithms rarely exist in a vacuum.
@@
 
-In Ivy one often wants to specify system-wide state-changes.
-For this reason Ivy allows iterator-free assignments involving all instances of a type.
+In IVy one often wants to specify system-wide state-changes.
+For this reason IVy allows iterator-free assignments involving all instances of a type.
 To make such an assignment one can use placeholders that are denoted by uppercase tokens.
@@
 This is mostly useful in combination with non-deterministic assignments that one wants to restrict.
-The following action chooses a random node to assign to \textit{y}, but will not allow control to pass if the chosen node has failed.
-Therefore the actions will choose a random non-failed node to send a message to x.
+The following action non-deterministically chooses a node to assign to \textit{y}, but will not allow control to pass if the chosen node has failed.
+Therefore the actions will choose a non-failed node to send a message to x.
 If there is no node that has not failed, this action cannot terminate.
@@
 \begin{lstlisting}
-action random_node_sends_message_to(x:node) = {
+action some_node_sends_message_to(x:node) = {
   y := *;
@@
 it is risky to use \textit{assume} for the same reason: one might dismiss a valid counterexample to the specification
-by assuming it cannot happen. To give an extreme example, putting $assume\ false$ at the beginning of an exported action will cause all of Ivy's checks for induction to come out positive.
+by assuming it cannot happen. To give an extreme example, putting $assume\ false$ at the beginning of an exported action will cause all of IVy's checks for induction to come out positive.
 
@@
 First order logic is undecidable in general, but still many interesting properties can be expressed in a subset of first oder logic called the ``decidable fragment''.
-Ivy's documentation defines this subset of first-order formulas.
+IVy's documentation defines this subset of first-order formulas.
 \cite{decid}
 
-For example, Ivy only allows stratified functions.
+For example, IVy only allows stratified functions.
 Meaning that a graph, with every type being a node and every function mapping from type $A$ to type $B$ be an edge from $A$ to $B$, has to be cycle free.
@@
 
-While a complete understanding of decidability in Ivy is advantageous, to use Ivy successfully it is not strictly necessary.
-This is because Ivy assists the user and tries to automatically determine if formulas lie outside of the decidable fragment.
+While a complete understanding of decidability in IVy is advantageous, to use IVy successfully it is not strictly necessary.
+This is because IVy assists the user and tries to automatically determine if formulas lie outside of a decidable fragment.
 It warns users about specifications that it estimates are undecidable.
 It must be noted that IVy's check is conservative as ``fully unfolding all of the definitions could be exponential''.
-Ivy may report specifications as undecidable which are in fact decidable.
+IVy may report specifications as undecidable which are in fact decidable.
 For this reason IVy allows the check to be deactivated.
@@
 
-The algorithm that was implemented in IVy as part of this work is called \textit{Floodset} and was taken out of Lynch's book ``Distributed Algorithms'' \cite{refNancy}.
+The algorithm that was verified in IVy as part of this work is called \textit{Floodset} and was taken out of Lynch's book ``Distributed Algorithms'' \cite{refNancy}.
 \textit{Floodset} is a consensus algorithm for synchronous networks with failing processes.
@@
 
-The network of processes is represented as a directed graph $G=(V,E)$. Where $n$ is $|V|$, the number of nodes in the network. Each node represents a process in the network.
+The network of processes is represented as a directed graph $G=(V,E)$. The number $|V|$ of nodes in the network is denoted by $n$. Each node represents a process in the network.
 The communication paths between nodes are represented by directed edges in the graph.
@@
 \textbf{Proof}:\\
-\textit{Indirect}: If there exists a value $v \in V$ that is in $W_i(r)$ but not in $W_j(r)$ after round $r$, in which no process failed, that implies, that some node $k$ sent a value to $i$ that it did not send to $j$, or $i$ failed to send a member of $W_i(r-1)$.
+\textit{Indirect}: Suppose no process failed during round $r$, and there exists a value $v \in V$ that is in $W_i(r)$ but not in $W_j(r)$.
+This implies, that some node $k$ sent a value to $i$ that it did not send to $j$, or $i$ failed to send a member of $W_i(r-1)$.
 Since all nodes broadcast their $W$, this is only possible if node $k$ or $i$ failed to send at least one message.
@@
 \noindent \textbf{Lemma 3}:\\
-After $f+1$ rounds there has been a round $r'$ in which no process failed\\
+After $f+1$ rounds there has been a round $r'$ in which no process failed.\\
 \textbf{Proof}:\\
@@
 
-\noindent The Lemmas allow proving the theorem:\\
+\noindent We now prove correctness of \textit{Floodset} by proving the three correctness conditions \textit{agreement}, \textit{validity}
+ and \textit{termination}. \\
 \noindent \textbf{Theorem}:\\
@@
 \textbf{Agreement}:\\
+\textbf{Proof}:\\
 From \textit{Lemma 4} we know what $W_i = W_j$ after $f+1$ rounds. Since all decision conditions are the same and only depend on $W$, agreement is satisfied.\\
 \textbf{Validity}:\\
+\textbf{Proof}:\\
 If all processes start with the same initial value $v \in V$, then there exists no value $v', v \neq v'$, that is part of any $W$.
@@
 \textbf{Termination}:\\
+\textbf{Proof}:\\
 Trivially, nodes decide after $f+1$ rounds.\\
@@
 To summarize: we have a parameterized system with $n$ process and $f$ number of faults. We want to show \textit{Floodset} is correct regardless of the parameters.
-It is trivial to verify for a small number of $n,\ f$, but the question is ``does \textit{Floodset} work for any assignment of $n,\ f,\ n < f$''.
+It is straightforward to check for a small number of $n,\ f$, but the question is ``does \textit{Floodset} work for any assignment of $n,\ f,\ n < f$''.
 To answer this question the algorithm was encoded in the IVy language and then inductive invariants were constructed to prove correctness.
@@
 The assumption $f<n$, ``number of failed nodes is bounded to an amount smaller than the total number of nodes'', can alternatively be expressed via $\exists n\ \neg failed(n)$.
-This is because the existence of a functioning node implies that the number of failed nodes $f$ is smaller than $n$.
+Intuitively, $f<n$ is true if and only if there is a process that has not failed.
+Or in other words, the existence of a functioning node implies that the number of failed nodes $f$ is smaller than $n$.
 
@@
 First the types, functions, relations and individual variables that are used throughout the program are presented.
-They are annotated with comments explaining their purpose.
-Their meaning will become more clear as the code is presented in the following paragraphs.
+
+\subsubsection{Types}
+The types used are $node$ and $value$.
+The set of processes is represented by $node$, $value$ represents the set of values $V$.
+Both types are uninterpreted, meaning they represent an arbitrary set of instances with at least one element.
+The cardinality of these sets is \textit{parameterized}, meaning IVy proves correctness for an arbitrary amount of $nodes$ and $values$.
+
+\subsubsection{Functions}
+The specification uses the function $val(N:node) : value$ to store the initial value of every node and $default(N:node) : value$ to represent each node's reference to the common default value.
+
+\subsubsection{Relations}
+The specification uses a variety of relations.
+E.g.
+to represent the set $W$, $w(N:node,V:value)$ is used.
+It specifies if a node $n$ has in its $W$ the value $v$.
+Another example is $message(N:node,V:value)$, which is used to specify the set of values any node is broadcasting in the current round.
+
+The purpose of the remaining relations can be found in the comments above their definition in the code below. Their meaning will become clear as the code is presented in the following paragraphs.
 
@@
 
-\noindent \textit{round} is the only exported action. It specifies the basic structure of Lynch's \textit{framework}.
+\noindent The only exported action is \textit{round}. It specifies the basic structure of Lynch's \textit{framework}.
 \noindent Line 2 makes it illegal for the environment to call \textit{round} if decisions already occurred.
@@
 
-The actual delivery of a message is modeled via the \\
-$incoming\_message(RECIPIENT:node,SENDER:node)$ relation. Where $incoming\_message(0,1) = true;\ incoming\_message(0,2) = true;$ means that node $0$ receives the messages from node $1$ and node $2$.
+The actual delivery of a message is modeled via the $incoming\_message(RECP:node,SEND:node)$ relation.
+Where $incoming\_message(0,1) = true;\ incoming\_message(0,2) = true;$ means that node $0$ receives the messages from node $1$ and node $2$.
 
 These relations allow us to use IVy's assignments to easily model \textit{Step 1} in a round of \textit{Floodset}.
-Line 2 assigns for all nodes $A$ and values $V$ to $message(A,V)$ the boolean value of $node.w(A,V) \land \neg failed(A)$.
+Line 2 assigns for all nodes $A$ and values $V$ to $message(A,V)$ the Boolean value of $node.w(A,V) \land \neg failed(A)$.
 In other words, every node that has not failed prepares a message with all values in its $w$.
-Line 3 states that all nodes that haven't failed send messages to all other nodes.
+Line 3 states that all nodes that have not failed send messages to all other nodes.
 
@@
 
-This section shows how the stopping failure model described by Lynch \cite{refNancy} and explained in \textit{section III ii.} was modeled in Ivy.
+This section shows how the stopping failure model described by Lynch \cite{refNancy} and explained in \textit{section III ii.} was modeled in IVy.
 
@@
 Lines 2 to 5 resemble the snippet of section \textit{vii.1}.
-In line 2 we choose a random number of nodes to fail.
+In line 2 we choose a non-determined number of nodes to fail.
 The \textit{assume} in line 3 then excludes all cases where a node that already failed was chosen to fail again.
@@
 However, truly verifying \textit{termination} would require checking that the system must reach the decision round, which means verifying  \textit{liveness}.
-Ivy theoretically allows \textit{liveness} checks by constructing a reduction of \textit{liveness} to \textit{safety} \cite{live} but this is out of the scope of this work.
-
-\subsection{Invariants for Inductiveness}
-If one were to run $ivy\_check$ on the program without specifying any
-additional invariants, IVy finds that the \textit{agreement} invariant is not inductive,
-and produces a counter example to induction (CTI). The CTI can be drawn in a GUI.
+IVy theoretically allows \textit{liveness} checks by constructing a reduction of \textit{liveness} to \textit{safety} \cite{live} but this is out of the scope of this work.
+
+\subsection{Invariants for Inductiveness and IVy's CTI-Viewer}
+If one were to run $ivy\_check$ on the program without specifying any additional invariants, IVy finds that neither the \textit{agreement} invariant, nor the \textit{validity} invariant is inductive.
+However, this does not immediately imply that \textit{Floodset}, or our specification of it, is incorrect.
+It means that the specified system can transition from a state in compliance with invariants to a state that is not.
+Thus, to prove \textit{Floodset} we need to find additional invariants that exclude unreachable states so that IVy can proof induction of the invariants that were found so far.
+
+To improve the readers' intuition for this process it helps to (wrongly) think of IVy as a system searching the state space for counterexamples to induction as follows:
+\begin{enumerate}
+  \item Choose an initial system state, i.e. assignments to types, relations and functions, that is compliant with all invariants.
+  \item Execute an exported action which mutates the initial system state.
+  \item Check if the system state after the execution violates any of the invariants, if yes, display counterexample to induction (CTI).
+  \item Goto 1.
+\end{enumerate}
+
+While this is not how IVy works internally, it illustrates an important point made in \textit{section I}: IVy does not consider the reachability of a system state.
+As long as a state is in compliance with the conjunction of all specified invariants it is assumed to be reachable.
+When IVy finds a counterexample to induction (CTI), it means that the system can transition from a state in compliance with invariants to a state that is not.
+It does not however imply, that there exists an execution path of exported actions so that the system, starting in its initial state defined in the \textit{init} action, can transition to a state that violates an invariant.
+We therefore have to strengthen our invariants, i.e.
+limit the state space that IVy can use to find CTIs, by adding invariants that exclude unreachable system states.
+
+As this is not a straightforward process, IVy uses a GUI to illustrate CTIs to the user.
+If, like in our case, multiple invariants fail the check for inductiveness, IVy draws a counterexample to the invariant that is defined earliest in the specification.
+For us, IVy shows a counterexample to \textit{agreement}.
 
@@
 to state 1 by executing \textit{round}. In the center IVy displays state 0, meaning the state
-before \textit{round} was executed. The brown edges specify the relation $w$, the purple edges
-show $val$.
-
-State 0 shows, that no node has their initial value, which is modeled by the relation $val$,
-in its $w$. The user must now realise, that this is an unreachable state, and exclude it by coming
-up with an invariant.
+before \textit{round} was executed. This state is in compliance with the set of invariants.
+The brown edges specify the relation $w$, the purple edges show $val$.
+Clicking on state 1 would show a state that violates the \textit{agreement} invariant.
+
+The user must now understand why state 0 is unreachable.
+Observing the graph of state 0, one can see that no node has their initial value, which is modeled by the relation $val$,
+in its $w$. The user must now exclude this state by coming up with an invariant.
 
@@
 
-The method for finding the remaining invariants is identical: Execute $ivy\_check$ $\rightarrow$ examine CTI $\rightarrow$ design invariant $\rightarrow$ execute $ivy\_check$.
+After adding this invariant, IVy can no longer construct a CTI with a state 0 that violates the added invariant.
+However, even after adding this invariant, IVy still fails to prove induction of \textit{agreement} and \textit{validity} as it still assumes a variety of unreachable states to be reachable.
+
+Thus, to prove \textit{Floodset}, the user must iterate the presented method for finding the remaining invariants: Execute $ivy\_check$ $\rightarrow$ examine CTI $\rightarrow$ design invariant $\rightarrow$ execute $ivy\_check$.
+
 The following snippets presents the necessary invariants for induction annotated with comments.
@@
 \begin{lstlisting}
+# initial value is in W
+invariant node.w(N, node.val(N))
+
 # the global default is the one the nodes know
@@
 
+The most complex of these invariants is on line 12, which is equivalent to Lemma 1 presented in \textit{section III}.
+The necessity to include such an invariant in the final proof shows that automatic proving in IVy still requires fundamental understanding of the algorithm.
+While the CTI-GUI helps users understand individual counterexamples, it does not come up with invariants on its own.
+It is therefore not a substitute for knowledge about the fundamental ideas behind an algorithm or protocol.
+
+After these four invariants are added to the specification, IVy can successfully prove induction of the system.
+As the specified invariants entail the correctness conditions, the correctness of \textit{Floodset} has been proven.
+
 \subsection{Checking framework correctness}
@@
 These invariants are not necessary for induction.
-Removing them will not cause any of Ivy's other checks to fail.
+Removing them will not cause any of IVy's other checks to fail.
 Their sole purpose is to give the user confidence that the \textit{framework} is specified correctly.
@@
 correctness. However, due to a variety of reasons, absolute certainty is never achieved.
-This section evaluates Ivy's usefulness pointing out strengths and limitations that Ivy users should keep in mind
+This section evaluates IVy's usefulness pointing out strengths and limitations that IVy users should keep in mind
 when working with the system.
@@
 One can not trivially determine, that the specification in the IVy language is logically equivalent to the actual algorithm the user wishes to verify.
-Whether the algorithm to be verified exists in the user's mind, in the form of pseudo-code or fully implemented in a programming language, no guarantee can be given in regards to the correctness of the translation of the algorithm into the Ivy language.
+Whether the algorithm to be verified exists in the user's mind, in the form of pseudo-code or fully implemented in a programming language, no guarantee can be given in regards to the correctness of the translation of the algorithm into the IVy language.
 Some examples illustrating this difficulty will be given in the following paragraphs.
 
-\subsubsection{Structural difference of Ivy and procedural languages}
+\subsubsection{Structural difference of IVy and procedural languages}
 The IVy language is structurally unlike classical procedural languages.
@@
 This comparison clearly shows the complex structural differences.
-While the pseudo-code is concerned only with state alteration of a single node, the Ivy language code simultaneously alters the state of all nodes.
+While the pseudo-code is concerned only with state alteration of a single node, the IVy language code simultaneously alters the state of all nodes.
 It uses an existential quantifier nested within a term.
@@
 As these two lines are so different, equivalence may not be obvious to a user.
-Proving equivalence of pseudo-code and Ivy language specification would be a non-trivial proof exercise that certainly trumps manually proving \textit{Floodset}'s correctness in difficulty.
+Proving equivalence of pseudo-code and IVy language specification would be a non-trivial proof exercise that certainly trumps manually proving \textit{Floodset}'s correctness in difficulty.
 
@@
 $assume$ statements may silently invalidate state configurations. They therefore pose a threat as they
-may allow false-positive Ivy checks.
+may allow false-positive IVy checks.
 To give a practical example, the author of this paper originally modeled the line
@@
 The negated part, $\gamma$, specifies that there exists a node $N$, that failed and failed to deliver a message to $J$ while successfully delivering a message to $K$, with $N,J,K$ distinct.
-By negating $\gamma$, we force Ivy to construct a counterexample such that $\gamma$ is satisfied.
+By negating $\gamma$, we force IVy to construct a counterexample such that $\gamma$ is satisfied.
 If IVy does not find a counterexample to this assertion, it means that $\gamma$ cannot occur.
@@
 The specification would therefore be erroneous.
-One has to finally remove the failing assertion to allow Ivy to prove induction again.
+One has to finally remove the failing assertion to allow IVy to prove induction again.
 
@@
 \subsection{Necessity of deep understanding}
-It has become apparent, that successful verification with IVy is not possible without deep knowledge about how the algorithm functions. Ivy is not a ``translate and go'' solution.
+It has become apparent, that successful verification with IVy is not possible without deep knowledge about how the algorithm functions. IVy is not a ``translate and go'' solution.
 Evidence of this include the complex invariant that requires knowledge about \textit{lemma 1} or the alternative encoding for $f<n$ and the decision condition $if\ rounds = f+1$.
@@
 \subsection{Algorithm failure in theory and practice}
-As a user it is important to understand that Ivy's goal is to verify algorithmic correctness on a high level.
+As a user it is important to understand that IVy's goal is to verify algorithmic correctness on a high level.
 It does not concern itself with implementation details.
@@
 It would have been no help to verify the Heartbeat protocol in IVy before implementing it in C, since the error was in an implementation detail allowing a buffer underflow.
-The Ivy user must realise, that the implementation of the algorithm must be tested through unit testing and/or bounded model checkers, despite successful verification in IVy.
-
-\subsection{Time of specification}
-Unfortunately the author missed the opportunity to scientifically record the time commitment of verifying \textit{Floodset} in Ivy.
-He estimates the time from starting to learn Ivy to successfully proving \textit{Floodset} at around 150 hours.
-In the case of \textit{Floodset} manual proving would have certainly been significantly faster.
-Drawing a general conclusion about practicability of Ivy is not possible from this data.
-
-For future work, it may be interesting to study whether specification in Ivy gets easier or harder than manual proving as algorithmic complexity increases.
+The IVy user must realise, that the implementation of the algorithm must be tested through unit testing and/or bounded model checkers, despite successful verification in IVy.
+
+\subsection{Time of specification and practicability}
+Unfortunately the author missed the opportunity to scientifically record the time commitment of verifying \textit{Floodset} in IVy.
+He estimates the time from starting to learn IVy to successfully proving \textit{Floodset} at around 150 hours.
+This surprisingly large number can be attributed to a variety of factors:
+\begin{itemize}
+  \item Willingness to write a modular specification utilizing iteration.
+  \item Attempting to count failed nodes through iteration.
+  \item Having to discover an alternative encoding to $f < n$.
+  \item Having to discover an alternative \textit{decision condition}.
+  \item The believe that a superficial understanding of the algorithm is enough, as the proof is ``automatic'', which in turn led to a considerable amount of time investment to discover the invariants.
+\end{itemize}
+While drawing general conclusions about the practicability of IVy from the authors work is not possible, an attempt to quantitatively evaluate on the practicability is made in the following paragraphs.
+The reader is urged to read these thoughts with an increased amount of scepticism.
+
+As \textit{Floodset} is a fairly simple algorithm, an expert in the area of distributed systems will likely be able to manually prove it in fewer hours than it would take them to design an adequate specification and discover inductive invariants.
+
+For future work, it would be interesting to study whether specification in IVy gets easier or harder than manual proving as algorithmic complexity increases.
+The author \textbf{believes} that an increase in complexity of the algorithm would increase the complexity of specification in the IVy language more than it would increase the complexity of manual proving.
+This believe is based on the observation, that both automatic and manual proofs require considerable effort to discover the fundamental proof ideas.
+However, automatic verification introduces the additional hurdle of specification - which may well require alternative encodings and workarounds more complex than in this paper.
+It can therefore be argued, that the work for automatic verification is a superset of the work of manual proving and therefore fundamentally more work-intensive.
+
+For this reason, the author does not see IVy in the role of a ``quick and easy'' alternative to manual proving.
+Instead, he finds it fits alongside manual proofs, as a high-cost, high-reward tool reusing the fundamental proof ideas in an automated fashion with the goal of increasing confidence in a proof's soundness.
+An example of this approach can be found in \cite{wrongProof}, where a bug in an algorithm was found using automated verification techniques, despite a published manual proof of correctness.
+
+Thus, while automatic verification is a time consuming task, it may still be useful to commit to it, as it increases confidence about algorithmic correctness alongside a manual proof.
 
@@
 \section{Conclusion}
-This paper has presented the Ivy system and important Ivy language features. It classified IVy
+This paper has presented the IVy system and important IVy language features. It classified IVy
 as part of a two step process of algorithm verification.
@@
     \bibitem{limits} K. Apt, D. Kozen ``Limits of Automatic Verification of Finite-State Concurrent Systems'', 1986, Available \textit{http://www.cs.cornell.edu/~kozen/Papers/LimitsForAutomatic.pdf}
-    \bibitem{ivy} O. Padon, K. McMillan, A. Panda, M. Safiv, S. Shoham, ``Ivy: Safety Verification by Interactive Generalization'', Available \textit{https://www.cs.tau.ac.il/~odedp/ivy.pdf}
+    \bibitem{ivy} O. Padon, K. McMillan, A. Panda, M. Safiv, S. Shoham, ``IVy: Safety Verification by Interactive Generalization'', Available \textit{https://www.cs.tau.ac.il/~odedp/ivy.pdf}
     \bibitem{live} O.
@@
     \bibitem{github} C. Claus, ``BA'', \textit{https://github.com/calvinclaus/ba}, 2018
+    \bibitem{wrongProof} P. Lincoln, J. Rushby. ``A formally verified algorithm for interactive consistency under a hybrid fault model.'', 1993
 

### 2018-08-22 paper/paper.tex
implemented comments3

@@
 Lines 2 to 5 resemble the snippet of section \textit{vii.1}.
-In line 2 we choose a non-determined number of nodes to fail.
+In line 2 we non-deterministically choose a number of nodes to fail.
 The \textit{assume} in line 3 then excludes all cases where a node that already failed was chosen to fail again.
@@
 
-\subsection{Equivalence of specification and algorithm}
+\subsection{Specification challenges}
+
+It has become apparent that correctly specifying an algorithm in IVy is not an easy feat.
+This sections summarizes and evaluates on the variety of challenges concerning correct specification that were mentioned in this paper.
+
+\subsubsection{Equivalence of specification and algorithm}
 One can not trivially determine, that the specification in the IVy language is logically equivalent to the actual algorithm the user wishes to verify.
 Whether the algorithm to be verified exists in the user's mind, in the form of pseudo-code or fully implemented in a programming language, no guarantee can be given in regards to the correctness of the translation of the algorithm into the IVy language.
-Some examples illustrating this difficulty will be given in the following paragraphs.
-
-\subsubsection{Structural difference of IVy and procedural languages}
+Some examples illustrating this difficulty will be given in the following paragraphs.\\
+
+\noindent \textbf{Structural difference of IVy and procedural languages}\\
 The IVy language is structurally unlike classical procedural languages.
@@
 Proving equivalence of pseudo-code and IVy language specification would be a non-trivial proof exercise that certainly trumps manually proving \textit{Floodset}'s correctness in difficulty.
-
-\subsubsection{Lack of modularity}
+\\
+
+\noindent \textbf{Lack of modularity}\\
 Furthermore, it has already been shown in section \textit{IV.
@@
 
-\subsection{Accidental invalidation of legal state configurations with \textit{assume}}
-$assume$ statements may silently invalidate state configurations. They therefore pose a threat as they
-may allow false-positive IVy checks.
-To give a practical example, the author of this paper originally modeled the line
-
-\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
-\begin{lstlisting}
-node.w(N,V) := (node.w(N,V) |
-               (~node.failed(N) &
-                exists J . (node.incoming_message(N,J) &
-                             node.message(J,V))));
-\end{lstlisting}
-\end{mdframed}
-
-\noindent as
-
-\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
-\begin{lstlisting}
-node.tmp_w(A,V) := node.w(A V);
-node.w(A,V) := *;
-assume node.w(A,V) <-> (node.tmp_w(A,V) |
-                        ((node.incoming_message(A,J) &
-                          node.message(J,V))));
-\end{lstlisting}
-\end{mdframed}
-The second snippet misses an existential quantifier on line 4, and therefore specifies behaviour that is unlike that of \textit{Floodset}.
-The author did not discover that this $assume$ is bogus and excludes valid system states from being verified.
-The error only became apparent when it was discovered that correctness could be proven in the absence of an invariant that should have been absolutely necessary for induction.
-
-\subsection{Miss-specification of system context}
+\subsubsection{Miss-specification of system context}
 As has been pointed out, it is not just the algorithm that must be correctly specified.
@@
 
-\subsection{Necessity of deep understanding}
-It has become apparent, that successful verification with IVy is not possible without deep knowledge about how the algorithm functions. IVy is not a ``translate and go'' solution.
-Evidence of this include the complex invariant that requires knowledge about \textit{lemma 1} or the alternative encoding for $f<n$ and the decision condition $if\ rounds = f+1$.
-
-\subsection{Algorithm failure in theory and practice}
-As a user it is important to understand that IVy's goal is to verify algorithmic correctness on a high level.
-It does not concern itself with implementation details.
-However, a variety of bugs and exploits do not stem from logically unsound algorithmic foundations but implementation errors.
-A wonderful example is the Heartbleed exploit in OpenSSL.
-It would have been no help to verify the Heartbeat protocol in IVy before implementing it in C, since the error was in an implementation detail allowing a buffer underflow.
-The IVy user must realise, that the implementation of the algorithm must be tested through unit testing and/or bounded model checkers, despite successful verification in IVy.
-
-\subsection{Time of specification and practicability}
+\subsubsection{Necessity of deep understanding}
+It has become apparent, that successful specification with IVy is not possible without deep knowledge about how the algorithm functions. IVy is not a ``translate and go'' solution.
+Evidence include the alternative encoding for $f<n$ and for the decision condition $if\ rounds = f+1$.
+
+\subsection{Verification challenges}
+
+The following sections summarizes and evaluates on the variety of challenges concerning verification that were mentioned in this paper.
+
+\subsubsection{Accidental invalidation of legal state configurations with \textit{assume}}
+$assume$ statements may silently invalidate state configurations. They therefore pose a threat as they
+may allow false-positive IVy checks.
+To give a practical example, the author of this paper originally modeled the line
+
+\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+node.w(N,V) := (node.w(N,V) |
+               (~node.failed(N) &
+                exists J . (node.incoming_message(N,J) &
+                             node.message(J,V))));
+\end{lstlisting}
+\end{mdframed}
+
+\noindent as
+
+\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
+\begin{lstlisting}
+node.tmp_w(A,V) := node.w(A V);
+node.w(A,V) := *;
+assume node.w(A,V) <-> (node.tmp_w(A,V) |
+                        ((node.incoming_message(A,J) &
+                          node.message(J,V))));
+\end{lstlisting}
+\end{mdframed}
+The second snippet misses an existential quantifier on line 4, and therefore specifies behaviour that is unlike that of \textit{Floodset}.
+The author did not discover that this $assume$ is bogus and excludes valid system states from being verified.
+The error only became apparent when it was discovered that correctness could be proven in the absence of an invariant that should have been absolutely necessary for induction.
+
+\subsubsection{Hard to discover inductive invariants}
+Successfully discovering inductive invariants for \textit{Floodset} required a flawless understanding of the principles underlying the algorithm.
+Evidence of this includes the complex invariant that requires knowledge about \textit{lemma 1}.
+Users must be prepared to gain strong foundational understanding of the reasons why an algorithm works before attempting to find inductive invariants for it.
+
+\subsubsection{Manual effort and practicability}
 Unfortunately the author missed the opportunity to scientifically record the time commitment of verifying \textit{Floodset} in IVy.
@@
 
+\subsubsection{Algorithm failure in theory and practice}
+As a user it is important to understand that IVy's goal is to verify algorithmic correctness on a high level.
+It does not concern itself with implementation details.
+However, a variety of bugs and exploits do not stem from logically unsound algorithmic foundations but implementation errors.
+A wonderful example is the Heartbleed exploit in OpenSSL.
+It would have been no help to verify the Heartbeat protocol in IVy before implementing it in C, since the error was in an implementation detail allowing a buffer underflow.
+The IVy user must realise, that the implementation of the algorithm must be tested through unit testing and/or bounded model checkers, despite successful verification in IVy.
+
 \subsection{User friendliness}
@@
 through individual lines and their effect on the state.
-
-Therefore, if a user is capable of navigating the limitations with care, IVy is absolutely adequate for proving the logical soundness of an algorithm.
 
@@
 
-It highlighted the systems strengths to be user friendliness, interactivity and a powerful logic based language that allows for
-modeling of complex system state transitions in simple one-liners.
-
-It concludes that IVy's user friendliness in presenting CTIs - so long as the user is vary of the presented limitations - enables
-finding inductive invariants for proving logical soundness of an algorithm relatively easily.
+It highlighted the systems strengths to be user friendliness, interactivity and a powerful logic based language that allows for modeling of complex system state transitions in simple one-liners.
+
+It concludes that IVy's user friendliness in presenting CTIs - so long as the user is vary of the presented limitations - enables finding inductive invariants for proving logical soundness of an algorithm relatively easily.
 
